---
title: "2_resazurin_calc_analysis"
author: "Samuel Gurr"
date: "2026-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
knitr::opts_knit$set(root.dir = "C:/Users/gurrs/Documents/6_Github_repositories/GurrLab/Predictive_phenomics/RAnalysis") # Sam's
# knitr::opts_knit$set(root.dir = "C:/ < Anna add here > /Predictive_phenomics/RAnalysis") # Anna's root dir on her computer
```


```{r}
library(readxl)
library(stringr)
library(pracma)
library(ggplot2)
library(ggpubr)
```

```{r}

# metadata
metadata_counts <- read.csv("data/Pacific_oyster/resazurin/metadata.csv", sep=',')
plate_notes     <- metadata_counts %>%  dplyr::select(c('PlateID','Well','Note','X')) %>% dplyr::rename(obs_volume = X)
families        <- read.csv("data/Pacific_oyster/YC26-ebv-crosses-broodpops.csv", sep=',') %>% dplyr::rename(Family = cross)

# outputs from 1_resazurin_raw_compiler
# life stages during larval development
rawdf_72hpf         <- read.csv("output/resazurin/raw/rawdf_72hpf.csv")
rawdf_15dpf         <- read.csv("output/resazurin/raw/rawdf_15dpf.csv")
rawdf_48dpf         <- read.csv("output/resazurin/raw/rawdf_48dpf.csv")
# temperature ramp experiment
rawdf_20260630_28C  <- read.csv("output/resazurin/raw/rawdf_20260630_28C.csv")

rawdf_20260702_30C  <- read.csv("output/resazurin/raw/rawdf_20260702_30C.csv")

rawdf_20260704_32C  <- read.csv("output/resazurin/raw/rawdf_20260704_32C.csv")

rawdf_20260706_34C  <- read.csv("output/resazurin/raw/rawdf_20260706_34C.csv")

rawdf_20260710_38C  <- read.csv("output/resazurin/raw/rawdf_20260710_38C.csv")

```



### no gain included in this function

* what to do for other timepoints with gain differences? run these manually - look below for how

```{r  AUC_calculate function}

AUC_calculate <- function (file_in) {

                output.df <- data.frame()
                loop.df_m   <- data.frame(unique(file_in$PlateID)) # call specific df
                loop.df_i   <- data.frame(unique(file_in$Well))  # call specific df
                # Call the cumulative dataframe that we will write to in the for loop below
                output.cumulative.df <- data.frame() # start dataframe 
                output.df            <- data.frame() # start dataframe 
                output.cor.df        <- data.frame() # start dataframe 
                running.df           <- data.frame(matrix(nrow = 1, ncol = 9)) # create dataframe to save cumunalitively during for loop
                colnames(running.df) <- c('Date', 'Treatment', 'PlateID', 'Age', 'Well', 'Family', 'Type', 'Count', 'AUC') # names for columns in the for loop
                          
                          for(m in 1:nrow(loop.df_m)){
                            plate_id      <- file_in %>%   # call specific df
                                                          dplyr::filter(PlateID %in% loop.df_m[m,1])
                                  for(i in 1:nrow(loop.df_i)){
                                    well.over.time       <- plate_id %>% 
                                                                  dplyr::filter(Well %in% loop.df_i[i,1]) %>% 
                                                                  dplyr::arrange(time) %>% 
                                                                  dplyr::mutate(hours = as.numeric(difftime(time, 
                                                                                                            time[1], 
                                                                                                            units = "hours")),
                                                                                delta_florescence = flourescence - flourescence[1],
                                                                                AUC = pracma::trapz(
                                                                                                    hours, 
                                                                                                    delta_florescence)) %>% 
                                                                  dplyr::select('Date', 'Treatment', 'PlateID', 'Age', 'time', 'Family', 'hours',
                                                                                'Well', 'Type', 'Count', 'AUC', 'delta_florescence')
                                                        
                                    output.cumulative.df <- rbind(output.cumulative.df, well.over.time)
                                    
                                    running.df$Date      <- well.over.time$Date[1]
                                    running.df$Treatment <- well.over.time$Treatment[1]
                                    running.df$PlateID   <- well.over.time$PlateID[1]
                                    running.df$Age       <- well.over.time$Age[1]
                                    running.df$Well      <- well.over.time$Well[1]
                                    running.df$Family    <- well.over.time$Family[1]
                                    running.df$Type      <- well.over.time$Type[1]
                                    running.df$Count     <- well.over.time$Count[1]
                                    running.df$AUC       <- well.over.time$AUC[1]
                                    
                                    output.df <- rbind(output.df,
                                                       running.df) #bind to a cumulative list dataframe
                                    # print(output.df) # print to monitor progress
                 
                                    } # wells within plate i end
      
                           } # wells within plate m end
            
            return(output.df)
            
} # function end
```

```{r AUC_calculate_TempRamp}

AUC_calculate_TempRamp <- function (file_in) {

                output.df            <- data.frame()
                outputBlanks.df      <- data.frame() # start dataframe 
                output.cumulative.df <- data.frame() # start dataframe 
                loop.df_m   <- data.frame(unique(file_in$gain))  # call specific df
                loop.df_i   <- data.frame(unique(file_in$Well))  # call specific df
                
                # Call the cumulative dataframe that we will write to in the for loop below
                running.df           <- data.frame(matrix(nrow = 1, ncol = 10)) # create dataframe to save cumunalitively during for loop
                colnames(running.df) <- c('Date', 'Treatment', 'PlateID', 'gain', 'Age', 'Well', 'Family', 'Type', 'Count', 'AUC') # names for columns in the for loop
                
                for(m in 1:nrow(loop.df_m)){
                  plate_id      <- file_in %>%   # call specific df !!! change to data_merged_seed34C <METADATA NEEDED FOR THIS!>
                                                dplyr::filter(gain %in% loop.df_m[m,1])
                  
                          for(i in 1:nrow(loop.df_i)){
                            plate_id$flourescence <- as.numeric(plate_id$flourescence)
                            # plate_id$Treatment    <- 34 # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            # plate_id$Age          <- "seed" # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            # plate_id$Family       <- NA # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            # plate_id$Type         <- NA # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            # plate_id$Count        <- NA # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            plate_id$gain         <- loop.df_m[m,1] # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            well.over.time        <- plate_id %>%
                              
                                                          dplyr::filter(Well %in% loop.df_i[i,1]) %>% 
                                                          dplyr::arrange(time) %>% 
                                                          dplyr::mutate(hours = as.numeric(difftime(time, 
                                                                                                    time[1], 
                                                                                                    units = "hours")),
                                                                        delta_florescence = flourescence - flourescence[1],
                                                                        AUC = pracma::trapz(
                                                                                            hours, 
                                                                                            delta_florescence)) #%>% 
                                                          # dplyr::select('Date', 'PlateID', 'gain', 'time', 'hours', 'Type'
                                                          #               'Well',  'AUC', 'delta_florescence')
                                                
                            output.cumulative.df <- rbind(output.cumulative.df, 
                                                          well.over.time)
                            

                            running.df$Date      <- well.over.time$Date[1]
                            running.df$Treatment <- well.over.time$Treatment[1]
                            running.df$PlateID   <- well.over.time$PlateID[1]
                            running.df$gain      <- well.over.time$gain[1]
                            running.df$Age       <- well.over.time$Age[1]
                            running.df$Well      <- well.over.time$Well[1]
                            running.df$Family    <- well.over.time$Family[1]
                            running.df$Type      <- well.over.time$Type[1]
                            running.df$Count     <- well.over.time$Count[1]
                            running.df$AUC       <- well.over.time$AUC[1]
                            
                            output.df <- rbind(output.df,
                                               (running.df %>% dplyr::filter(Type %in% 'Sample'))
                                               )  #bind to a cumulative list dataframe %>% 

                            outputBlanks.df <- rbind(outputBlanks.df,
                                               (running.df %>% dplyr::filter(Type %in% c('blank', 'shell')))
                                               )  

                            #                         )
                          } # for(i in 1:nrow(loop.df_i)){

                } # for(m in 1:nrow(loop.df_m)){
            mean.AUC.blanks <- outputBlanks.df %>% 
                                     dplyr::group_by(Type, gain) %>% 
                                     dplyr::summarise(
                                             meanAUC = mean(AUC),
                                             n = dplyr::n())
                
            return(list(DeltaFlo.df = output.cumulative.df,
                        AUC.df      = output.df, 
                        AUCblanks   = mean.AUC.blanks
                        ))
}


```


```{r run AUC_calculate function}

# run function AUC_calculate
AUC_calculate_72hpf  <- AUC_calculate(rawdf_72hpf)
AUC_calculate_15dpf  <- AUC_calculate(rawdf_15dpf)
AUC_calculate_48dpf  <- AUC_calculate(rawdf_48dpf)

# run function AUC
AUC_calculate_20260630_28C  <- AUC_calculate_TempRamp(rawdf_20260630_28C)
AUC_calculate_20260702_30C  <- AUC_calculate_TempRamp(rawdf_20260702_30C)
AUC_calculate_20260704_32C  <- AUC_calculate_TempRamp(rawdf_20260704_32C) 
AUC_calculate_20260706_34C  <- AUC_calculate_TempRamp(rawdf_20260706_34C)
AUC_calculate_20260710_38C  <- AUC_calculate_TempRamp(rawdf_20260710_38C) 
```




## Plot data 

```{r Plot data}

ggplot(data = AUC_calculate_20260630_28C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260702_30C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260704_32C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260706_34C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260710_38C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)
```



## Clean up data 


#### 72 hpf data 
  * temperatures 26 and 32 (only 2!)
  * normalized by count (# of individuals counted per well, fixed in formalin)
  * true blanks are instances where 'Sample' Count == 0, ignore 'Blank' for "Type" column (only here!)
```{r 72 hpf clean}

# dataset == AUC_calculate_72hpf


# SAMPLES  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_72hpf.Samples <- AUC_calculate_72hpf %>% dplyr::filter(Type %in% 'Sample' & Count != 0)
#  raw plot to view well relationship with count
ggplot(data =AUC_calculate_72hpf.Samples, aes(x=Count, y = AUC)) +
       geom_point() + 
       labs(title = "AUC (not blank corrected) by Count") +
       facet_wrap(~Treatment)
# verdict: Yes, there is a slight, expected, relatoinship with increase AUC with more indivs/well
# raw plot to view well bias (potential) 
ggplot(data =AUC_calculate_72hpf.Samples, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for samples by well") +
       facet_wrap(~Treatment)
# verdict: NO DISCERNABLE WELL BIAS


# BLANKS  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_72hpf.Blanks  <- AUC_calculate_72hpf %>% dplyr::filter(Type %in% 'Sample' & Count == 0) 
# raw plot to view well bias (potential) 

ggplot(data =AUC_calculate_72hpf.Blanks, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for blanks by well") 
# verdict:  NO DISCERNABLE WELL BIAS


# MEAN BLANKS BY PLATE + TREATMENT  :::::::::::::::::::::::::::::::::::::::
# data
MeanBlanks_72hpf <- AUC_calculate_72hpf.Blanks %>% 
                        dplyr::group_by(PlateID) %>% 
                        dplyr::summarise(
                          meanAUC = mean(AUC)
                        )


# MERGE BLANKS WITH SAMPLES, CALCULATE AUC_COR AND AUC_ADJ  ::::::::::::::
AUC_calculate_72hpf <- merge(AUC_calculate_72hpf.Samples, MeanBlanks_72hpf, by = "PlateID") %>%  # merge by plateID
                                            dplyr::mutate(
                                              AUC.cor = AUC - meanAUC,
                                              AUC.cor.norm = AUC.cor / Count,
                                              Family_EBV = dplyr::case_when(Family > 50 ~ 'Low',
                                                              Family < 20 ~ 'High')
                                            ) %>% 
                                            dplyr::filter(AUC.cor >= 0) %>% # omits negative, AUC samples less than blank
                                            dplyr::select(
                                              "Date",
                                              "Age",
                                              "Treatment",
                                              "PlateID",
                                              "Well",
                                              "Family",
                                              "Family_EBV",
                                              "Count",
                                              "AUC",
                                              "AUC.cor",
                                              "AUC.cor.norm"
                                            )
  

# PLOTTING ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# AUC corrected adn normalized, all data by family and treatment
ggplot(AUC_calculate_72hpf, aes(x = as.factor(Family), y = AUC.cor.norm)) + 
  geom_point() + facet_wrap(~Treatment)


# means SE data, for pub 
AUC_calculate_72hpf.PLOT <- ggplot(data =AUC_calculate_72hpf, 
                                   aes(x=as.factor(Family), 
                                       y = AUC.cor.norm)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 72 hpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)

```


#### 15 dpf data 
  * temperatures 26, 30, 34, and 38C
  * normalized by count (# of individuals counted per well, fixed in formalin)
  * true blanks are denoted as 'Blank' in the "Type" column
```{r 15 dpf clean}

# dataset == AUC_calculate_15dpf
unique(AUC_calculate_15dpf$Treatment)


# SAMPLES  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_15dpf.Samples <- AUC_calculate_15dpf %>% dplyr::filter(Type %in% 'Sample' & Count != 0)
#  raw plot to view well relationship with count
ggplot(data =AUC_calculate_15dpf.Samples, aes(x=Count, y = AUC)) +
       geom_point() + 
       labs(title = "AUC (not blank corrected) by Count") +
       facet_wrap(~Treatment)
# verdict: Yes, there is a slight, expected, relationship with increase AUC with more indivs/well
# raw plot to view well bias (potential) 
ggplot(data =AUC_calculate_15dpf.Samples, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for samples by well") +
       facet_wrap(~Treatment)
# verdict: NO DISCERNABLE WELL BIAS


# BLANKS  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_15dpf.Blanks  <- AUC_calculate_15dpf %>% dplyr::filter(Type %in% 'Blank') 
# raw plot to view well bias (potential) 

ggplot(data =AUC_calculate_15dpf.Blanks, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for blanks by well") +
       facet_wrap(~Treatment)
  
# verdict:  NO DISCERNABLE WELL BIAS, BUT OMIT THOSE >1000 OUTLIERS
boxplot.stats(AUC_calculate_15dpf.Blanks$AUC)$out

AUC_calculate_15dpf.Blanks.OM <- AUC_calculate_15dpf.Blanks %>% dplyr::filter(!AUC > 1000)

# MEAN BLANKS BY PLATE + TREATMENT  :::::::::::::::::::::::::::::::::::::::
# data
MeanBlanks_15dpf <- AUC_calculate_15dpf.Blanks.OM %>% 
                        dplyr::group_by(PlateID) %>% 
                        dplyr::summarise(
                          meanAUC = mean(AUC)
                        )


# MERGE BLANKS WITH SAMPLES, CALCULATE AUC_COR AND AUC_ADJ  ::::::::::::::
AUC_calculate_15dpf <- merge(AUC_calculate_15dpf.Samples, MeanBlanks_15dpf, by = "PlateID") %>%  # merge by plateID
                                            dplyr::mutate(
                                              AUC.cor = AUC - meanAUC,
                                              AUC.cor.norm = AUC.cor / Count,
                                              Family_EBV = dplyr::case_when(Family > 50 ~ 'Low',
                                                              Family < 20 ~ 'High')
                                              ) %>% 
                                            dplyr::filter(AUC.cor >= 0) %>% # omits negative, AUC samples less than blank
                                            dplyr::select(
                                              "Date",
                                              "Age",
                                              "Treatment",
                                              "PlateID",
                                              "Well",
                                              "Family",
                                              "Family_EBV",
                                              "Count",
                                              "AUC",
                                              "AUC.cor",
                                              "AUC.cor.norm"
                                            )

```

```{r 15 dpf clean plotting}
# PLOTTING ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# AUC corrected adn normalized, all data by family and treatment
ggplot(AUC_calculate_15dpf, aes(x = as.factor(Family), y = AUC.cor.norm)) + 
  geom_point() + facet_wrap(~Treatment)


# means SE data, for pub 
AUC_calculate_15dpf.PLOT <- ggplot(data =AUC_calculate_15dpf, 
                                   aes(x=as.factor(Family), 
                                       y = AUC.cor.norm)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 15 dpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_15dpf.PLOT2 <- ggplot(data =AUC_calculate_15dpf, 
                                   aes(x=as.factor(Family_EBV), 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 15 dpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_15dpf.PLOT3a <- ggplot(data =AUC_calculate_15dpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 15 dpf"
                                    )  +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() 

AUC_calculate_15dpf.PLOT3b <- ggplot(data =AUC_calculate_15dpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 15 dpf"
                                    )  +
                                    geom_smooth(method = "loess", se = FALSE) +
                                    theme_classic() + 
                                    facet_wrap(~Family_EBV)

AUC_calculate_15dpf.PLOT4 <- ggplot(AUC_calculate_15dpf, 
                                    aes(x = Treatment, 
                                        y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                      stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() + 
                                    facet_wrap(~Family)


```


#### 48 dpf data 
  * temperatures 26, 30, 34, and 38C
  * normalized by count (# of individuals counted per well, fixed in formalin)
  * true blanks are denoted as 'Blank' in the "Type" column
```{r 48 dpf clean}
# dataset == AUC_calculate_48dpf
unique(AUC_calculate_48dpf$Treatment)


# SAMPLES  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_48dpf.Samples <- AUC_calculate_48dpf %>% dplyr::filter(Type %in% 'Sample' & Count != 0)
# raw plot to view well bias (potential) 
ggplot(data =AUC_calculate_48dpf.Samples, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for samples by well") +
       facet_wrap(~Treatment)
# verdict: NO DISCERNABLE WELL BIAS


# BLANKS  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_48dpf.Blanks  <- AUC_calculate_48dpf %>% dplyr::filter(Type %in% 'Blank') 
# raw plot to view well bias (potential) 

ggplot(data =AUC_calculate_48dpf.Blanks, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for blanks for shell (1) and no shell (NA)") +
       facet_wrap(~Count)
# verdict:  YES, blanks with an empty shell included are higher!

# MEAN BLANKS BY PLATE + TREATMENT  :::::::::::::::::::::::::::::::::::::::
# data
MeanBlanks_48dpf <- AUC_calculate_48dpf.Blanks %>% 
                        dplyr::group_by(PlateID, Count) %>% 
                        dplyr::summarise(
                          meanAUC = mean(AUC)
                        ) %>% 
                        tidyr::pivot_wider(
                          names_from = Count,
                          values_from = meanAUC
                        ) %>% 
                        dplyr::rename(
                          meanAUC.shell  = "1",
                          meanAUC.blank  = "NA"
                        )


# MERGE BLANKS WITH SAMPLES, CALCULATE AUC_COR AND AUC_ADJ  ::::::::::::::
AUC_calculate_48dpf <- merge(AUC_calculate_48dpf.Samples, MeanBlanks_48dpf, by = "PlateID") %>%  # merge by plateID
                                            dplyr::mutate(
                                              AUC.cor = AUC - meanAUC.blank,
                                              AUC.corSHELL = AUC - meanAUC.shell,
                                              AUC.cor.norm = AUC.corSHELL / Count,
                                              Family_EBV = dplyr::case_when(Family > 50 ~ 'Low',
                                                              Family < 20 ~ 'High')
                                              ) %>% 
                                            dplyr::filter(AUC.corSHELL >= 0) %>% # omits negative, AUC samples less than blank
                                            dplyr::select(
                                              "Date",
                                              "Age",
                                              "Treatment",
                                              "PlateID",
                                              "Well",
                                              "Family",
                                              "Family_EBV",
                                              "Count",
                                              "AUC",
                                              "AUC.cor",
                                              "AUC.cor.norm"
                                            )

```

```{r 48 dpf clean plotting}
# PLOTTING ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# AUC corrected adn normalized, all data by family and treatment
ggplot(AUC_calculate_48dpf, aes(x = as.factor(Family), y = AUC.cor.norm)) + 
  geom_point() + facet_wrap(~Treatment)


# means SE data, for pub 
AUC_calculate_48dpf.PLOT <- ggplot(data =AUC_calculate_48dpf, 
                                   aes(x=as.factor(Family), 
                                       y = AUC.cor.norm)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 48 dpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_48dpf.PLOT2 <- ggplot(data =AUC_calculate_48dpf, 
                                   aes(x=as.factor(Family_EBV), 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 48 dpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_48dpf.PLOT3a <- ggplot(data =AUC_calculate_48dpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 48 dpf"
                                    )  +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() 


AUC_calculate_48dpf.PLOT3b <- ggplot(data =AUC_calculate_48dpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 48 dpf"
                                    )  +
                                    geom_smooth(method = "loess", se = FALSE) +
                                    theme_classic() + 
                                    facet_wrap(~Family_EBV)

AUC_calculate_48dpf.PLOT4 <- ggplot(AUC_calculate_48dpf, 
                                    aes(x = Treatment, 
                                        y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                      stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0
                                    ) +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() + 
                                    facet_wrap(~Family)


```

```{r Output CalculatedAUC larvae and early seed}

Master_larvae_seed <- rbind(AUC_calculate_72hpf,
                            AUC_calculate_15dpf,
                            AUC_calculate_48dpf)


write.csv(Master_larvae_seed, "output/resazurin/calculated/CalulatedAUC_Master_larvae_earlyseed.csv")



```

```{r}

ggarrange(
  AUC_calculate_15dpf.PLOT3a,AUC_calculate_15dpf.PLOT3b,
  AUC_calculate_48dpf.PLOT3a, AUC_calculate_48dpf.PLOT3b,
  nrow=2,
  ncol=2
)

```


## Temperature Ramp


#### Master file


ADJUSTING FOR GAIN, NORMALIZE TO 65
  * Going from lower gain to higher
          AUC * 0.145 for every 1 unit increase in gain 
  * Going from higher gain to lower
          AUC * 0.095 for every 1 unit decrease in gain
          
          
```{r}

0.095*(77-65)
AUC_calculate_20260630_28C$AUC.df %>% 
            dplyr::filter(gain == 77) %>% 
            dplyr::mutate(AUC.BlankShell = (AUC_calculate_20260630_28C$AUCblanks %>% dplyr::filter(gain == 77 & Type %in% 'shell'))$meanAUC,
                          AUC.cor = AUC - AUC.BlankShell
            )
                          
AUC_calculate_20260702_30C.FINAL <- merge(AUC_calculate_20260702_30C$AUC.df,
                                    (as.data.frame(AUC_calculate_20260702_30C$AUCblanks) %>% dplyr::filter(Type %in% 'shell') %>% dplyr::select('gain', 'meanAUC')),
                                    by = 'gain') %>% 
                                          dplyr::rename(AUC.BlankShell = meanAUC) %>% 
                                          dplyr::mutate(AUC.cor = AUC - AUC.BlankShell) %>% 
                                          dplyr::filter(gain == 65)
                                  
                     
AUC_calculate_20260704_32C.FINAL <- merge(AUC_calculate_20260704_32C$AUC.df,
                                    (as.data.frame(AUC_calculate_20260704_32C$AUCblanks) %>% dplyr::filter(Type %in% 'shell') %>% dplyr::select('gain', 'meanAUC')),
                                    by = 'gain') %>% 
                                          dplyr::rename(AUC.BlankShell = meanAUC) %>% 
                                          dplyr::mutate(AUC.cor = AUC - AUC.BlankShell)


AUC_calculate_20260706_34C.FINAL <- merge(AUC_calculate_20260706_34C$AUC.df,
                                    (as.data.frame(AUC_calculate_20260706_34C$AUCblanks) %>% dplyr::filter(Type %in% 'shell') %>% dplyr::select('gain', 'meanAUC')),
                                    by = 'gain') %>% 
                                          dplyr::rename(AUC.BlankShell = meanAUC) %>% 
                                          dplyr::mutate(AUC.cor = AUC - AUC.BlankShell) %>% 
                                          dplyr::filter(gain == 65)



0.145 * (65-60)
AUC_calculate_20260710_38C.FINAL <- merge(AUC_calculate_20260710_38C$AUC.df,
                                    (as.data.frame(AUC_calculate_20260710_38C$AUCblanks) %>% dplyr::filter(Type %in% 'shell') %>% dplyr::select('gain', 'meanAUC')),
                                    by = 'gain') %>% 
                                          dplyr::rename(AUC.BlankShell = meanAUC) %>% 
                                          dplyr::mutate(AUC.cor = (AUC - AUC.BlankShell)*1.725) # convert to 65 gain


```

```{r  Output CalculatedAUC juvenile TempRamp}
Master <- rbind(AUC_calculate_20260702_30C.FINAL,
      AUC_calculate_20260704_32C.FINAL,
      AUC_calculate_20260706_34C.FINAL,
      AUC_calculate_20260710_38C.FINAL)

write.csv(Master, "output/resazurin/calculated/CalulatedAUC_Master_TempRamp.csv")



```


```{r plotting juvenile TempRamp}
Master_meansFamily <- Master %>% 
                        dplyr::group_by(Date, Treatment, Family, Age) %>% 
                        dplyr::summarise(meanAUC = mean(AUC.cor)) %>% 
                        dplyr::mutate(Family_EBV = dplyr::case_when(Family >= 48 ~ 'Low',
                                                                    Family <= 20 ~ 'High'))

ggplot(data = Master_meansFamily,
       aes(x = Treatment,
           y = meanAUC,
           color = Family_EBV,
           group = Family)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
                                    theme_classic() 


 ggplot(data =Master_meansFamily, 
                                   aes(x=Treatment, 
                                       y = meanAUC, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, correct to blank shell)",
                                      title = "C. magallana juvniles temp ramp"
                                    )  +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() 
```


```{r stats by family juvenile TempRamp}

AOV30C.Family <- aov(
                      lm(
                        AUC.cor ~ Family,
                         data = (Master %>% dplyr::filter(Treatment %in% "30C"))))
shapiro.test(resid(AOV30C.Family)) # 0.003699 - N.S,
kruskal.test(AUC.cor ~ Family,data = (Master %>% dplyr::filter(Treatment %in% "30C"))) # 0.007907 - significant

1AOV32C.Family <- aov(
                      lm(
                        AUC.cor ~ Family,
                         data = (Master %>% dplyr::filter(Treatment %in% "32C"))))
shapiro.test(resid(AOV32C.Family)) # 0.02742 - N.S,
kruskal.test(AUC.cor ~ Family,data = (Master %>% dplyr::filter(Treatment %in% "32C"))) # 0.04516 - significant

AOV34C.Family <- aov(
                      lm(
                        AUC.cor ~ Family,
                         data = (Master %>% dplyr::filter(Treatment %in% "34C"))))
shapiro.test(resid(AOV34C.Family)) # 0.004611 - N.S,
kruskal.test(AUC.cor ~ Family,data = (Master %>% dplyr::filter(Treatment %in% "34C"))) # 0.003453 - significant


AOV38C.Family <- aov(
                      lm(
                        AUC.cor ~ Family,
                         data = (Master %>% dplyr::filter(Treatment %in% "38C"))))
shapiro.test(resid(AOV38C.Family)) # 0.008765 - N.S,
kruskal.test(AUC.cor ~ Family,data = (Master %>% dplyr::filter(Treatment %in% "38C"))) # 0.1768 - NOT SIGNIFICANT

```

```{r stats by high low EBV groups juvenile TempRamp}

AOV30C.highvlow <- aov(
                      lm(
                        meanAUC ~ Family_EBV,
                         data = (Master_meansFamily %>% dplyr::filter(Treatment %in% "30C"))))
shapiro.test(resid(AOV30C.highvlow)) # 0.4541
summary(AOV30C.highvlow) # 0.0727 .

AOV32C.highvlow <- aov(
                      lm(
                        meanAUC ~ Family_EBV,
                         data = (Master_meansFamily %>% dplyr::filter(Treatment %in% "32C"))))
shapiro.test(resid(AOV32C.highvlow)) # 0.09189
summary(AOV32C.highvlow) # 0.00217 **.

AOV34C.highvlow <- aov(
                      lm(
                        meanAUC ~ Family_EBV,
                         data = (Master_meansFamily %>% dplyr::filter(Treatment %in% "34C"))))
shapiro.test(resid(AOV34C.highvlow)) # 0.2049
summary(AOV34C.highvlow) # 0.144

AOV38C.highvlow <- aov(
                      lm(
                        meanAUC ~ Family_EBV,
                         data = (Master_meansFamily %>% dplyr::filter(Treatment %in% "38C"))))
shapiro.test(resid(AOV38C.highvlow)) # 0.3482
summary(AOV38C.highvlow) # 0.845
```
