---
title: "2_resazurin_calc_analysis"
author: "Samuel Gurr"
date: "2026-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
knitr::opts_knit$set(root.dir = "C:/Users/gurrs/Documents/6_Github_repositories/GurrLab/Predictive_phenomics/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "~/Documents/Predictive_phenomics/RAnalysis") # Anna's root dir on her computer
```


```{r}
library(readxl)
library(stringr)
library(pracma)
library(ggplot2)
library(ggpubr)
```

```{r load raw}

# metadata
metadata_counts <- read.csv("data/Pacific_oyster/resazurin/metadata.csv", sep=',')
plate_notes     <- metadata_counts %>%  dplyr::select(c('PlateID','Well','Note','X')) %>% dplyr::rename(obs_volume = X)
families        <- read.csv("data/Pacific_oyster/YC26-ebv-crosses-broodpops.csv", sep=',') %>% dplyr::rename(Family = cross)

# outputs from 1_resazurin_raw_compiler
# life stages during larval development
rawdf_72hpf         <- read.csv("output/resazurin/raw/rawdf_72hpf.csv") %>%  dplyr::filter(!PlateID %in% 'plate_1_low') %>% dplyr::filter(!Note %in% c('has no volume', 'has no volume ', 'has low volume', 'has less volume'))
rawdf_15dpf         <- read.csv("output/resazurin/raw/rawdf_15dpf.csv")
rawdf_48dpf         <- read.csv("output/resazurin/raw/rawdf_48dpf.csv")
# temperature ramp experiment
rawdf_20260622_20C  <- read.csv("output/resazurin/raw/rawdf_20260622_20C.csv")

rawdf_20260626_24C  <- read.csv("output/resazurin/raw/rawdf_20260626_24C.csv")

rawdf_20260628_26C  <- read.csv("output/resazurin/raw/rawdf_20260628_26C.csv")

rawdf_20260630_28C  <- read.csv("output/resazurin/raw/rawdf_20260630_28C.csv")

rawdf_20260704_32C  <- read.csv("output/resazurin/raw/rawdf_20260704_32C.csv")

rawdf_20260702_30C  <- read.csv("output/resazurin/raw/rawdf_20260702_30C.csv")

rawdf_20260706_34C  <- read.csv("output/resazurin/raw/rawdf_20260706_34C.csv")

rawdf_20260710_38C  <- read.csv("output/resazurin/raw/rawdf_20260710_38C.csv")

```


```{r data adj}
adjdf_72hpf <- rawdf_72hpf %>% dplyr::filter(!PlateID %in% 'plate_1_low',
                                             !Note %in% c('has no volume', 'has no volume ', 'has low volume', 'has less volume'),
                                             !Count %in% 0)

```



### no gain included in this function

* what to do for other timepoints with gain differences? run these manually - look below for how

```{r  AUC_calculate function}

AUC_calculate <- function (file_in) {

                output.df <- data.frame()
                loop.df_m   <- data.frame(unique(file_in$PlateID)) # call specific df
                loop.df_i   <- data.frame(unique(file_in$Well))  # call specific df
                # Call the cumulative dataframe that we will write to in the for loop below
                output.cumulative.df <- data.frame() # start dataframe 
                output.df            <- data.frame() # start dataframe 
                output.cor.df        <- data.frame() # start dataframe 
                running.df           <- data.frame(matrix(nrow = 1, ncol = 9)) # create dataframe to save cumunalitively during for loop
                colnames(running.df) <- c('Date', 'Treatment', 'PlateID', 'Age', 'Well', 'Family', 'Type', 'Count', 'AUC') # names for columns in the for loop
                          
                          for(m in 1:nrow(loop.df_m)){
                            plate_id      <- file_in %>%   # call specific df
                                                          dplyr::filter(PlateID %in% loop.df_m[m,1])
                                  for(i in 1:nrow(loop.df_i)){
                                    well.over.time       <- plate_id %>% 
                                                                  dplyr::filter(Well %in% loop.df_i[i,1]) %>% 
                                                                  dplyr::arrange(time) %>% 
                                                                  dplyr::mutate(hours = as.numeric(difftime(time, 
                                                                                                            time[1], 
                                                                                                            units = "hours")),
                                                                                delta_florescence = flourescence - flourescence[1],
                                                                                AUC = pracma::trapz(
                                                                                                    hours, 
                                                                                                    delta_florescence)) %>% 
                                                                  dplyr::select('Date', 'Treatment', 'PlateID', 'Age', 'time', 'Family', 'hours',
                                                                                'Well', 'Type', 'Count', 'AUC', 'delta_florescence')
                                                        
                                    output.cumulative.df <- rbind(output.cumulative.df, well.over.time)
                                    
                                    running.df$Date      <- well.over.time$Date[1]
                                    running.df$Treatment <- well.over.time$Treatment[1]
                                    running.df$PlateID   <- well.over.time$PlateID[1]
                                    running.df$Age       <- well.over.time$Age[1]
                                    running.df$Well      <- well.over.time$Well[1]
                                    running.df$Family    <- well.over.time$Family[1]
                                    running.df$Type      <- well.over.time$Type[1]
                                    running.df$Count     <- well.over.time$Count[1]
                                    running.df$AUC       <- well.over.time$AUC[1]
                                    
                                    output.df <- rbind(output.df,
                                                       running.df) #bind to a cumulative list dataframe
                                    # print(output.df) # print to monitor progress
                 
                                    } # wells within plate i end
      
                           } # wells within plate m end
            
            return(list(DeltaFlo.df = output.cumulative.df,
                        AUC.df      = output.df
                        ))            
} # function end
```

```{r AUC_calculate_TempRamp}

AUC_calculate_TempRamp <- function (file_in) {

                output.df            <- data.frame()
                outputBlanks.df      <- data.frame() # start dataframe 
                output.cumulative.df <- data.frame() # start dataframe 
                loop.df_m   <- data.frame(unique(file_in$gain))  # call specific df
                loop.df_i   <- data.frame(unique(file_in$Well))  # call specific df
                
                # Call the cumulative dataframe that we will write to in the for loop below
                running.df           <- data.frame(matrix(nrow = 1, ncol = 10)) # create dataframe to save cumunalitively during for loop
                colnames(running.df) <- c('Date', 'Treatment', 'PlateID', 'gain', 'Age', 'Well', 'Family', 'Type', 'Count', 'AUC') # names for columns in the for loop
                
                for(m in 1:nrow(loop.df_m)){
                  plate_id      <- file_in %>%   # call specific df !!! change to data_merged_seed34C <METADATA NEEDED FOR THIS!>
                                                dplyr::filter(gain %in% loop.df_m[m,1])
                  
                          for(i in 1:nrow(loop.df_i)){
                            plate_id$flourescence <- as.numeric(plate_id$flourescence)
                            # plate_id$Treatment    <- 34 # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            # plate_id$Age          <- "seed" # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            # plate_id$Family       <- NA # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            # plate_id$Type         <- NA # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            # plate_id$Count        <- NA # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            plate_id$gain         <- loop.df_m[m,1] # manually insert the temperature WAITING FOR METADATA TO CALL plate_id IN 'm'
                            well.over.time        <- plate_id %>%
                              
                                                          dplyr::filter(Well %in% loop.df_i[i,1]) %>% 
                                                          dplyr::arrange(time) %>% 
                                                          dplyr::mutate(hours = as.numeric(difftime(time, 
                                                                                                    time[1], 
                                                                                                    units = "hours")),
                                                                        delta_florescence = flourescence - flourescence[1],
                                                                        AUC = pracma::trapz(
                                                                                            hours, 
                                                                                            delta_florescence)) #%>% 
                                                          # dplyr::select('Date', 'PlateID', 'gain', 'time', 'hours', 'Type'
                                                          #               'Well',  'AUC', 'delta_florescence')
                                                
                            output.cumulative.df <- rbind(output.cumulative.df, 
                                                          well.over.time)
                            

                            running.df$Date      <- well.over.time$Date[1]
                            running.df$Treatment <- well.over.time$Treatment[1]
                            running.df$PlateID   <- well.over.time$PlateID[1]
                            running.df$gain      <- well.over.time$gain[1]
                            running.df$Age       <- well.over.time$Age[1]
                            running.df$Well      <- well.over.time$Well[1]
                            running.df$Family    <- well.over.time$Family[1]
                            running.df$Type      <- well.over.time$Type[1]
                            running.df$Count     <- well.over.time$Count[1]
                            running.df$AUC       <- well.over.time$AUC[1]
                            
                            output.df <- rbind(output.df,
                                               (running.df %>% dplyr::filter(Type %in% 'Sample'))
                                               )  #bind to a cumulative list dataframe %>% 

                            outputBlanks.df <- rbind(outputBlanks.df,
                                               (running.df %>% dplyr::filter(Type %in% c('blank', 'shell')))
                                               )  

                            #                         )
                          } # for(i in 1:nrow(loop.df_i)){

                } # for(m in 1:nrow(loop.df_m)){
            mean.AUC.blanks <- outputBlanks.df %>% 
                                     dplyr::group_by(Type, gain) %>% 
                                     dplyr::summarise(
                                             meanAUC = mean(AUC),
                                             n = dplyr::n())
                
            return(list(DeltaFlo.df = output.cumulative.df,
                        AUC.df      = output.df, 
                        AUCblanks   = mean.AUC.blanks
                        ))
}


```


```{r run AUC_calculate function}

# run function AUC_calculate
AUC_calculate_72hpf  <- AUC_calculate(adjdf_72hpf) # omit plate 1 low 
AUC_calculate_15dpf  <- AUC_calculate(rawdf_15dpf)
AUC_calculate_48dpf  <- AUC_calculate(rawdf_48dpf)

# run function AUC
AUC_calculate_20260622_20C  <- AUC_calculate_TempRamp(rawdf_20260622_20C)
AUC_calculate_20260626_24C  <- AUC_calculate_TempRamp(rawdf_20260626_24C)
AUC_calculate_20260628_26C  <- AUC_calculate_TempRamp(rawdf_20260628_26C)
AUC_calculate_20260630_28C  <- AUC_calculate_TempRamp(rawdf_20260630_28C)
AUC_calculate_20260702_30C  <- AUC_calculate_TempRamp(rawdf_20260702_30C)
AUC_calculate_20260704_32C  <- AUC_calculate_TempRamp(rawdf_20260704_32C) 
AUC_calculate_20260706_34C  <- AUC_calculate_TempRamp(rawdf_20260706_34C)
AUC_calculate_20260710_38C  <- AUC_calculate_TempRamp(rawdf_20260710_38C) 

AUC_calculate_72hpf

AUC_calculate_20260622_20C
```


```{r view outliers}

# 72 hpf

x.df72_plate_1_high <- ((AUC_calculate_72hpf$DeltaFlo.df %>%  dplyr::filter(PlateID %in% 'plate_1_high') %>%  dplyr::filter(Type %in% 'Blank'))$AUC)
boxplot.stats(x.df72_plate_1_high)$out # ouliter = 1731.65 AUC


x.df72_plate_2_high <- ((AUC_calculate_72hpf$DeltaFlo.df %>%  dplyr::filter(PlateID %in% 'plate_2_high') %>%  dplyr::filter(Type %in% 'Blank'))$AUC)
boxplot.stats(x.df72_plate_2_high)$out# outlier AUC = 1927.892

x.df72_plate_2_low <- ((AUC_calculate_72hpf$DeltaFlo.df %>%  dplyr::filter(PlateID %in% 'plate_2_low') %>%  dplyr::filter(Type %in% 'Blank'))$AUC)
boxplot.stats(x.df72_plate_2_low)$out # no outliers

# omit the two outliers from both dataframes 
AUC_calculate_72hpf$DeltaFlo.df <- (AUC_calculate_72hpf$DeltaFlo.df) %>% dplyr::filter(!(Type %in% 'Blank' & AUC >= 1731.650))
AUC_calculate_72hpf$AUC.df <- (AUC_calculate_72hpf$AUC.df) %>% dplyr::filter(!(Type %in% 'Blank' & AUC >= 1731.650))
# View((AUC_calculate_72hpf$DeltaFlo.df))
```


## Plot data 

```{r Plot data}


ggplot(data = (AUC_calculate_72hpf$DeltaFlo.df),
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~PlateID)


ggplot(data = AUC_calculate_20260622_20C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260626_24C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260628_26C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260630_28C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260702_30C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260704_32C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260706_34C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)

ggplot(data = AUC_calculate_20260710_38C$DeltaFlo.df,
    aes(x = time,
        y = delta_florescence,
        group = Well,
        color = Type)) +
    geom_point() +
    geom_line() +
  facet_wrap(~gain)
```



## Clean up data, Plot, Run Stats


#### 72 hpf data 
  * temperatures 26 and 32 (only 2!)
  * normalized by count (# of individuals counted per well, fixed in formalin)
  * true blanks are denoted as 'Blank' in the "Type" column
  * CONCLUSION: Sig effect at 32C showing reduced metabolic activity for low EBV families (77% lower than High EBVs)
```{r 72 hpf clean}

# dataset == AUC_calculate_72hpf


# SAMPLES  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_72hpf.Samples <- AUC_calculate_72hpf$AUC.df %>% dplyr::filter(Type %in% 'Sample')
nrow(AUC_calculate_72hpf$AUC.df )
nrow(AUC_calculate_72hpf.Samples)
#  raw plot to view well relationship with count
ggplot(data =AUC_calculate_72hpf.Samples, aes(x=Count, y = AUC)) +
       geom_point() + 
       labs(title = "AUC (not blank corrected) by Count") +
       facet_wrap(~Treatment)
# verdict: Yes, there is a slight, expected, relatoinship with increase AUC with more indivs/well
# raw plot to view well bias (potential) 
ggplot(data =AUC_calculate_72hpf.Samples, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for samples by well") +
       facet_wrap(~Treatment)
# verdict: NO DISCERNABLE WELL BIAS


# BLANKS  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_72hpf.Blanks  <-  AUC_calculate_72hpf$AUC.df %>% dplyr::filter(Type %in% 'Blank') 
# raw plot to view well bias (potential) 

ggplot(data =AUC_calculate_72hpf.Blanks, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for blanks by well") 
# verdict:  NO DISCERNABLE WELL BIAS


# MEAN BLANKS BY PLATE + TREATMENT  :::::::::::::::::::::::::::::::::::::::
# data
MeanBlanks_72hpf <- AUC_calculate_72hpf.Blanks %>% 
                        dplyr::group_by(PlateID) %>% 
                        dplyr::summarise(
                          meanBlank.AUC = mean(AUC)
                        )


# MERGE BLANKS WITH SAMPLES, CALCULATE AUC_COR AND AUC_ADJ  ::::::::::::::
AUC_calculate_72hpf <- merge(AUC_calculate_72hpf.Samples, MeanBlanks_72hpf, by = "PlateID") %>%  # merge by plateID
                                            dplyr::mutate(
                                              AUC.cor = AUC - meanBlank.AUC,
                                              AUC.cor.norm = AUC.cor / Count,
                                              Family_EBV = dplyr::case_when(Family > 50 ~ 'Low',
                                                              Family < 20 ~ 'High')
                                            ) %>% 
                                            dplyr::filter(AUC.cor >= 0) %>% # omits negative, AUC samples less than blank
                                            dplyr::select(
                                              "Date",
                                              "Age",
                                              "Treatment",
                                              "PlateID",
                                              "Well",
                                              "Family",
                                              "Family_EBV",
                                              "Count",
                                              "meanBlank.AUC",
                                              "AUC",
                                              "AUC.cor",
                                              "AUC.cor.norm"
                                            )
  
nrow(AUC_calculate_72hpf)
nrow(AUC_calculate_72hpf.Samples)


```

```{r 72 hpf clean plotting}
# PLOTTING ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# AUC corrected adn normalized, all data by family and treatment
ggplot(AUC_calculate_72hpf, aes(x = as.factor(Family), y = AUC.cor.norm)) + 
  geom_point() + facet_wrap(~Treatment)


# means SE data, for pub 
AUC_calculate_72hpf.PLOT <- ggplot(data =AUC_calculate_72hpf, 
                                   aes(x=as.factor(Family), 
                                       y = AUC.cor.norm)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "C. magallana larvae @ 72 hpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)

AUC_calculate_72hpf.PLOT2 <- ggplot(data =AUC_calculate_72hpf, 
                                   aes(x=as.factor(Family_EBV), 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 72 hpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_72hpf.PLOT3a <- ggplot(data =AUC_calculate_72hpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    geom_jitter() +                                   
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 72 hpf"
                                    )  +
                                    # geom_smooth(method = "loess",               
                                    #             se = TRUE,                      
                                    #             level = 0.95,
                                    #             span = 0.8)  +
                                    theme_classic() 

AUC_calculate_72hpf.PLOT3b <- ggplot(data =AUC_calculate_72hpfMEANS, 
                                   aes(x=as.factor(Treatment), 
                                       y = as.numeric(meanAUC.cor.norm), 
                                        group =  as.factor(Family), 
                                        color = Family_EBV)) +
                                    geom_point() +
                                    # stat_summary(
                                    #   fun = mean,
                                    #   geom = "point",
                                    #   size = 3
                                    # ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 72 hpf"
                                    )  +
                                     geom_smooth(method = "lm", se = FALSE) +
                                    theme_classic() + 
                                    facet_wrap(~Family_EBV)

AUC_calculate_72hpf.PLOT4 <- ggplot(AUC_calculate_72hpf, 
                                    aes(x = Treatment, 
                                        y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                      stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() + 
                                    facet_wrap(~Family)

AUC_calculate_72hpfMEANS <- AUC_calculate_72hpf %>% group_by(Family, Treatment, Family_EBV) %>% dplyr::summarise(meanAUC.cor.norm = mean(AUC.cor.norm))



dodge_width <- 0.70



AUC_calculate_72hpfMEANS$Family_EBV <- factor(
  AUC_calculate_72hpfMEANS$Family_EBV,
  levels = c("Low", "High")
)

      
Plot_72hpf_MEAN.SE <- ggplot(
                            AUC_calculate_72hpfMEANS,
                            aes(
                              x = Treatment,
                              y = meanAUC.cor.norm,
                              colour = Family_EBV,
                              fill = Family_EBV,
                              group = Family_EBV
                            )
                          ) +
                            geom_point(
                              shape = 21,
                              size = 2.4,
                              alpha = 0.55,
                              position = position_jitterdodge(
                                jitter.width = 0.12,
                                dodge.width = dodge_width,
                                seed = 123
                              )
                            ) +
                            scale_fill_manual(values = c('#0072B2', '#E69F00')) +
                            stat_summary(
                              fun = mean,
                              geom = "point",
                              shape = 21,
                              size = 4,
                              colour = "black",
                              position = position_dodge(width = dodge_width)
                            ) +
                            stat_summary(
                              fun.data = mean_se,
                              geom = "errorbar",
                              width = 0.12,
                              linewidth = 0.8,
                              colour = "black",
                              position = position_dodge(width = dodge_width)
                            ) +
                            scale_colour_manual(values = c('#0072B2', '#E69F00')) +
                            labs(
                              title = "72 hpf",
                              x = "Temperature treatment",
                              y = "AUC (per individual)",
                              colour = "Independent variable 2",
                              fill = "Independent variable 2"
                            ) +
                            theme_classic() +
                            theme(legend.position = "none") 
      


```

```{r stats for 72 hpf}

# prepare subset to run t tests within temperature
unique(AUC_calculate_72hpfMEANS$Treatment) # "26C" "32C"
AUC_calculate_72hpfMEANS_26C <- AUC_calculate_72hpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '26C') %>% 
                                      dplyr::mutate(
                                        Family_EBV_num =
                                        dplyr::case_when(Family_EBV %in% 'Low' ~ 1,
                                                  Family_EBV %in% 'High' ~ 2,
                                                  )
                                      )

AUC_calculate_72hpfMEANS_32C <- AUC_calculate_72hpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '32C') %>% 
                                      dplyr::mutate(
                                        Family_EBV_num =
                                        dplyr::case_when(Family_EBV %in% 'Low' ~ 1,
                                                  Family_EBV %in% 'High' ~ 2,
                                                  )
                                      )
# MODEL ASSUMPTIONS for t test
# shapiro wilk to check for normality 
by(AUC_calculate_72hpfMEANS_26C$meanAUC.cor.norm, AUC_calculate_72hpfMEANS_26C$Family_EBV, shapiro.test) # 0.5897 - normal
by(AUC_calculate_72hpfMEANS_32C$meanAUC.cor.norm, AUC_calculate_72hpfMEANS_32C$Family_EBV, shapiro.test) # 0.2207 - normal

# levenes test to check for homogeneity of variance 
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_72hpfMEANS_26C) # 0.7377 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_72hpfMEANS_32C) # 0.04428 - fail

# conclusion
# run the Welch's t test for 26C, run the non parametric Wilcox rnak sum for 32C
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_72hpfMEANS_26C, var.equal = FALSE) # 0.6151 not significant
wilcox.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_72hpfMEANS_32C) # 0.008658

# run follow up of sig effect
AUC_calculate_72hpfMEANS_32C %>% dplyr::group_by(Family_EBV) %>%  dplyr::summarise(mean = mean(meanAUC.cor.norm))
# High	12.914278			
# Low	  2.942732
((12.914278-2.942732)/12.914278)*100 # 77% lower metabolic activity on average for the low EBV than High EBV fams

```


#### 15 dpf data 
  * temperatures 26, 30, 34, and 38C
  * normalized by count (# of individuals counted per well, fixed in formalin)
  * true blanks are denoted as 'Blank' in the "Type" column
  * CONCLUSION: No effect
```{r 15 dpf clean}

# dataset == AUC_calculate_15dpf
unique(AUC_calculate_15dpf$Treatment)


# SAMPLES  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_15dpf.Samples <- AUC_calculate_15dpf %>% dplyr::filter(Type %in% 'Sample' & Count != 0)
#  raw plot to view well relationship with count
ggplot(data =AUC_calculate_15dpf.Samples, aes(x=Count, y = AUC)) +
       geom_point() + 
       labs(title = "AUC (not blank corrected) by Count") +
       facet_wrap(~Treatment)
# verdict: Yes, there is a slight, expected, relationship with increase AUC with more indivs/well
# raw plot to view well bias (potential) 
ggplot(data =AUC_calculate_15dpf.Samples, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for samples by well") +
       facet_wrap(~Treatment)
# verdict: NO DISCERNABLE WELL BIAS


# BLANKS  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_15dpf.Blanks  <- AUC_calculate_15dpf %>% dplyr::filter(Type %in% 'Blank') 
# raw plot to view well bias (potential) 

ggplot(data =AUC_calculate_15dpf.Blanks, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for blanks by well") +
       facet_wrap(~Treatment)
  
# verdict:  NO DISCERNABLE WELL BIAS, BUT OMIT THOSE >1000 OUTLIERS
boxplot.stats(AUC_calculate_15dpf.Blanks$AUC)$out

AUC_calculate_15dpf.Blanks.OM <- AUC_calculate_15dpf.Blanks %>% dplyr::filter(!AUC > 1000)

# MEAN BLANKS BY PLATE + TREATMENT  :::::::::::::::::::::::::::::::::::::::
# data
MeanBlanks_15dpf <- AUC_calculate_15dpf.Blanks.OM %>% 
                        dplyr::group_by(PlateID) %>% 
                        dplyr::summarise(
                          meanAUC = mean(AUC)
                        )


# MERGE BLANKS WITH SAMPLES, CALCULATE AUC_COR AND AUC_ADJ  ::::::::::::::
AUC_calculate_15dpf <- merge(AUC_calculate_15dpf.Samples, MeanBlanks_15dpf, by = "PlateID") %>%  # merge by plateID
                                            dplyr::mutate(
                                              AUC.cor = AUC - meanAUC,
                                              AUC.cor.norm = AUC.cor / Count,
                                              Family_EBV = dplyr::case_when(Family > 50 ~ 'Low',
                                                              Family < 20 ~ 'High')
                                              ) %>% 
                                            dplyr::filter(AUC.cor >= 0) %>% # omits negative, AUC samples less than blank
                                            dplyr::select(
                                              "Date",
                                              "Age",
                                              "Treatment",
                                              "PlateID",
                                              "Well",
                                              "Family",
                                              "Family_EBV",
                                              "Count",
                                              "AUC",
                                              "AUC.cor",
                                              "AUC.cor.norm"
                                            )

```

```{r 15 dpf clean plotting}
# PLOTTING ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# AUC corrected adn normalized, all data by family and treatment
ggplot(AUC_calculate_15dpf, aes(x = as.factor(Family), y = AUC.cor.norm)) + 
  geom_point() + facet_wrap(~Treatment)


# means SE data, for pub 
AUC_calculate_15dpf.PLOT <- ggplot(data =AUC_calculate_15dpf, 
                                   aes(x=as.factor(Family), 
                                       y = AUC.cor.norm)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 15 dpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_15dpf.PLOT2 <- ggplot(data =AUC_calculate_15dpf, 
                                   aes(x=as.factor(Family_EBV), 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 15 dpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_15dpf.PLOT3a <- ggplot(data =AUC_calculate_15dpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 15 dpf"
                                    )  +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() 

AUC_calculate_15dpf.PLOT3b <- ggplot(data =AUC_calculate_15dpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 15 dpf"
                                    )  +
                                    geom_smooth(method = "loess", se = FALSE) +
                                    theme_classic() + 
                                    facet_wrap(~Family_EBV)

AUC_calculate_15dpf.PLOT4 <- ggplot(AUC_calculate_15dpf, 
                                    aes(x = Treatment, 
                                        y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                      stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() + 
                                    facet_wrap(~Family)



AUC_calculate_15dpfMEANS <- AUC_calculate_15dpf %>% group_by(Family, Treatment, Family_EBV) %>% dplyr::summarise(meanAUC.cor.norm = mean(AUC.cor.norm))



dodge_width <- 0.70



AUC_calculate_15dpfMEANS$Family_EBV <- factor(
  AUC_calculate_15dpfMEANS$Family_EBV,
  levels = c("Low", "High")
)
      
Plot_15dpf_MEAN.SE <- ggplot(
                            AUC_calculate_15dpfMEANS,
                            aes(
                              x = Treatment,
                              y = meanAUC.cor.norm,
                              colour = Family_EBV,
                              fill = Family_EBV,
                              group = Family_EBV
                            )
                          ) +
                            geom_point(
                              shape = 21,
                              size = 2.4,
                              alpha = 0.55,
                              position = position_jitterdodge(
                                jitter.width = 0.12,
                                dodge.width = dodge_width,
                                seed = 123
                              )
                            ) +
                            scale_fill_manual(values = c('#0072B2', '#E69F00')) +
                            stat_summary(
                              fun = mean,
                              geom = "point",
                              shape = 21,
                              size = 4,
                              colour = "black",
                              position = position_dodge(width = dodge_width)
                            ) +
                            stat_summary(
                              fun.data = mean_se,
                              geom = "errorbar",
                              width = 0.12,
                              linewidth = 0.8,
                              colour = "black",
                              position = position_dodge(width = dodge_width)
                            ) +
                            scale_colour_manual(values = c('#0072B2', '#E69F00')) +
                            labs(
                              title = "15 dpf",
                              x = "Temperature treatment",
                              y = "AUC (per individual)",
                              colour = "Independent variable 2",
                              fill = "Independent variable 2"
                            ) +
                            theme_classic() +
                            theme(legend.position = "none") 

```

```{r stats for 15 dpf}

#  T TEST ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# test for differences between  high and low EBV families WITHIN each temperature incubation

# prepare subset to run t tests within temperature
unique(AUC_calculate_15dpfMEANS$Treatment) # "26C" "30C" "34C" "38C"
AUC_calculate_15dpfMEANS_26C <- AUC_calculate_15dpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '26C') 
                                      
AUC_calculate_15dpfMEANS_30C <- AUC_calculate_15dpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '30C')

AUC_calculate_15dpfMEANS_34C <- AUC_calculate_15dpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '34C') 

AUC_calculate_15dpfMEANS_38C <- AUC_calculate_15dpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '38C')

# MODEL ASSUMPTIONS for t test
# shapiro wilk to check for normality 
by(AUC_calculate_15dpfMEANS_26C$meanAUC.cor.norm, AUC_calculate_15dpfMEANS_26C$Family_EBV, shapiro.test) # 0.1179 - normal
by(AUC_calculate_15dpfMEANS_30C$meanAUC.cor.norm, AUC_calculate_15dpfMEANS_30C$Family_EBV, shapiro.test) # 0.5064 - normal
by(AUC_calculate_15dpfMEANS_34C$meanAUC.cor.norm, AUC_calculate_15dpfMEANS_34C$Family_EBV, shapiro.test) # 0.727 - normal
by(AUC_calculate_15dpfMEANS_38C$meanAUC.cor.norm, AUC_calculate_15dpfMEANS_38C$Family_EBV, shapiro.test) # 0.09452 - normal

# levenes test to check for homogeneity of variance 
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_15dpfMEANS_26C) # 0.8619 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_15dpfMEANS_30C) # 0.7986 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_15dpfMEANS_34C) # 0.8392 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_15dpfMEANS_38C) # 0.5776 - pass

# conclusion
# all models for the four temperatures pass! Move forward with t test
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_15dpfMEANS_26C, var.equal = FALSE) # 0.5054 not significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_15dpfMEANS_30C, var.equal = FALSE) # 0.6647 not significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_15dpfMEANS_34C, var.equal = FALSE) # 0.9376 not significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_15dpfMEANS_38C, var.equal = FALSE) # 0.2013 not significant


#  TWO WAY ANOVA  ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# run two way anpva for EBV * temperature

TwoWayANOVA_15dpf <- aov(lm(meanAUC.cor.norm ~ Family_EBV*Treatment, data = AUC_calculate_15dpfMEANS))
# check model assumptions 
shapiro.test(resid(TwoWayANOVA_15dpf)) # 0.03088 - fail 
car::leveneTest(TwoWayANOVA_15dpf) # 0.9421 - pass
# homogeneity of variance failed, run the non parametric Shrier Ray Hare 
rcompanion::scheirerRayHare(
                            meanAUC.cor.norm ~ Family_EBV * Treatment,
                            data = AUC_calculate_15dpfMEANS
                          ) # no significant differences
#                      Df Sum Sq      H p.value
# Family_EBV            1    5.3 0.0272 0.86898
# Treatment             3  775.5 3.9566 0.26619
# Family_EBV:Treatment  3  332.2 1.6947 0.63811
# Residuals            40 8099.0  
```


#### 48 dpf data 
  * temperatures 26, 30, 34, and 38C
  * normalized by shell area measured in ImageJ
  * true blanks are denoted as 'Blank' in the "Type" column
  * CONCLUSION: Overall sig effect (two way anova) showing reduced metabolic activity for low EBV families (40% lower than High EBVs)
                Sig effects of EBV at 26C and 38 (Welch's t test), showing 65% lower at 26C and 50% lower at 38C for low EBV fams than High EBV fams
```{r 48 dpf clean}
# dataset == AUC_calculate_48dpf
unique((AUC_calculate_48dpf$AUC.df)$Treatment)


metadata_48dpf <- metadata_counts %>% dplyr::filter(Age %in% '48dpf') %>% dplyr::select(Family, Treatment, Well, Area_ImageJ)



# SAMPLES  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_48dpf.Samples <- AUC_calculate_48dpf$AUC.df %>% dplyr::filter(Type %in% 'Sample' & Count != 0)
nrow(AUC_calculate_48dpf.Samples) # 336
AUC_calculate_48dpf.Samples.merge <- merge(metadata_48dpf, AUC_calculate_48dpf.Samples)

# raw plot to view well bias (potential) 
ggplot(data =AUC_calculate_48dpf.Samples.merge, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for samples by well") +
       facet_wrap(~Treatment)
# verdict: NO DISCERNABLE WELL BIAS


# BLANKS  :::::::::::::::::::::::::::::::::::::::
# data
AUC_calculate_48dpf.Blanks  <- AUC_calculate_48dpf$AUC.df %>% dplyr::filter(Type %in% 'Blank') 
# raw plot to view well bias (potential) 

ggplot(data =AUC_calculate_48dpf.Blanks, aes(x=Well, y = AUC)) +
       geom_point() + 
       labs(title = "View the AUC for blanks for shell (1) and no shell (NA)") +
       facet_wrap(~Count)
# verdict:  YES, blanks with an empty shell included are higher!

# MEAN BLANKS BY PLATE + TREATMENT  :::::::::::::::::::::::::::::::::::::::
# data
MeanBlanks_48dpf <- AUC_calculate_48dpf.Blanks %>% 
                        dplyr::group_by(PlateID, Count) %>% 
                        dplyr::summarise(
                          meanAUC = mean(AUC)
                        ) %>% 
                        tidyr::pivot_wider(
                          names_from = Count,
                          values_from = meanAUC
                        ) %>% 
                        dplyr::rename(
                          meanAUC.shell  = "1",
                          meanAUC.blank  = "NA"
                        )


# MERGE BLANKS WITH SAMPLES, CALCULATE AUC_COR AND AUC_ADJ  ::::::::::::::
AUC_calculate_48dpf <- merge(AUC_calculate_48dpf.Samples.merge, MeanBlanks_48dpf, by = "PlateID") %>%  # merge by plateID
                                            dplyr::mutate(
                                              AUC.cor = AUC - meanAUC.blank,
                                              AUC.corSHELL = AUC - meanAUC.shell,
                                              AUC.cor.norm = AUC.corSHELL / Area_ImageJ, # divided by sahell area,
                                              Family_EBV = dplyr::case_when(Family > 50 ~ 'Low',
                                                              Family < 20 ~ 'High')
                                              ) %>% 
                                            dplyr::filter(AUC.corSHELL >= 0) %>% # omits negative, AUC samples less than blank
                                            dplyr::select(
                                              "Date",
                                              "Age",
                                              "Treatment",
                                              "PlateID",
                                              "Well",
                                              "Family",
                                              "Family_EBV",
                                              "Count",
                                              "AUC",
                                              "AUC.cor",
                                              "AUC.cor.norm"
                                            )

```

```{r 48 dpf clean plotting}
# PLOTTING ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# AUC corrected adn normalized, all data by family and treatment
ggplot(AUC_calculate_48dpf, aes(x = as.factor(Family), y = AUC.cor.norm)) + 
  geom_point() + facet_wrap(~Treatment)


# means SE data, for pub 
AUC_calculate_48dpf.PLOT <- ggplot(data =AUC_calculate_48dpf, 
                                   aes(x=as.factor(Family), 
                                       y = AUC.cor.norm)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 48 dpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_48dpf.PLOT2 <- ggplot(data =AUC_calculate_48dpf, 
                                   aes(x=as.factor(Family_EBV), 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 48 dpf"
                                    ) +
                                    theme_classic() +
                                    facet_wrap(~Treatment)


AUC_calculate_48dpf.PLOT3a <- ggplot(data =AUC_calculate_48dpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 48 dpf"
                                    )  +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() 


AUC_calculate_48dpf.PLOT3b <- ggplot(data =AUC_calculate_48dpf, 
                                   aes(x=Treatment, 
                                       y = AUC.cor.norm, 
                                        group =  Family, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, normalized to count)",
                                      title = "M. gigas larvae @ 48 dpf"
                                    )  +
                                    geom_smooth(method = "loess", se = FALSE) +
                                    theme_classic() + 
                                    facet_wrap(~Family_EBV)

AUC_calculate_48dpf.PLOT4 <- ggplot(AUC_calculate_48dpf, 
                                    aes(x = Treatment, 
                                        y = AUC.cor.norm, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                      stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0
                                    ) +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() + 
                                    facet_wrap(~Family)


AUC_calculate_48dpfMEANS <- AUC_calculate_48dpf %>% group_by(Family, Treatment, Family_EBV) %>% dplyr::summarise(meanAUC.cor.norm = mean(AUC.cor.norm))



dodge_width <- 0.70



AUC_calculate_48dpfMEANS$Family_EBV <- factor(
  AUC_calculate_48dpfMEANS$Family_EBV,
  levels = c("Low", "High")
)
      
Plot_48dpf_MEAN.SE <- ggplot(
                            AUC_calculate_48dpfMEANS,
                            aes(
                              x = Treatment,
                              y = meanAUC.cor.norm,
                              colour = Family_EBV,
                              fill = Family_EBV,
                              group = Family_EBV
                            )
                          ) +
                            geom_point(
                              shape = 21,
                              size = 2.4,
                              alpha = 0.55,
                              position = position_jitterdodge(
                                jitter.width = 0.12,
                                dodge.width = dodge_width,
                                seed = 123
                              )
                            ) +
                            scale_fill_manual(values = c('#0072B2', '#E69F00')) +
                            stat_summary(
                              fun = mean,
                              geom = "point",
                              shape = 21,
                              size = 4,
                              colour = "black",
                              position = position_dodge(width = dodge_width)
                            ) +
                            stat_summary(
                              fun.data = mean_se,
                              geom = "errorbar",
                              width = 0.12,
                              linewidth = 0.8,
                              colour = "black",
                              position = position_dodge(width = dodge_width)
                            ) +
                            scale_colour_manual(values = c('#0072B2', '#E69F00')) +
                            labs(
                              title = "48 dpf",
                              x = "Temperature treatment",
                              y = "AUC (individual shell area)",
                              colour = "Independent variable 2",
                              fill = "Independent variable 2"
                            ) +
                            theme_classic() +
                            theme(legend.position = "none") 
```

```{r stats for 48 dpf}

#  T TEST ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# test for differences between  high and low EBV families WITHIN each temperature incubation

# prepare subset to run t tests within temperature
unique(AUC_calculate_48dpfMEANS$Treatment) # "26C" "30C" "34C" "38C"
AUC_calculate_48dpfMEANS_26C <- AUC_calculate_48dpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '26C') 
                                      
AUC_calculate_48dpfMEANS_30C <- AUC_calculate_48dpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '30C')

AUC_calculate_48dpfMEANS_34C <- AUC_calculate_48dpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '34C') 

AUC_calculate_48dpfMEANS_38C <- AUC_calculate_48dpfMEANS %>% 
                                      dplyr::filter(Treatment %in% '38C')

# MODEL ASSUMPTIONS for t test
# shapiro wilk to check for normality 
by(AUC_calculate_48dpfMEANS_26C$meanAUC.cor.norm, AUC_calculate_48dpfMEANS_26C$Family_EBV, shapiro.test) # 0.2368 - normal
by(AUC_calculate_48dpfMEANS_30C$meanAUC.cor.norm, AUC_calculate_48dpfMEANS_30C$Family_EBV, shapiro.test) # 0.4337 - normal
by(AUC_calculate_48dpfMEANS_34C$meanAUC.cor.norm, AUC_calculate_48dpfMEANS_34C$Family_EBV, shapiro.test) # 0.1136 - normal
by(AUC_calculate_48dpfMEANS_38C$meanAUC.cor.norm, AUC_calculate_48dpfMEANS_38C$Family_EBV, shapiro.test) # 0.5243 - normal

# levenes test to check for homogeneity of variance 
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_48dpfMEANS_26C) # 0.06179 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_48dpfMEANS_30C) # 0.931 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_48dpfMEANS_34C) # 0.8691 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_48dpfMEANS_38C) # 0.2721 - pass

# conclusion
# all models for the four temperatures pass! Move forward with t test
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_48dpfMEANS_26C, var.equal = FALSE) # 0.03899  significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_48dpfMEANS_30C, var.equal = FALSE) # 0.7285 not significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_48dpfMEANS_34C, var.equal = FALSE) # 0.126 not significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = AUC_calculate_48dpfMEANS_38C, var.equal = FALSE) # 0.01043  significant

#26C followup
AUC_calculate_48dpfMEANS_26C %>% dplyr::group_by(Family_EBV) %>%  dplyr::summarise(mean = mean(meanAUC.cor.norm))
# High	586.6778			
# Low	  202.7337	
((586.6778-202.7337)/586.6778)*100 # 65% lower metabolic activity on average for the low EBV than High EBV fams


#38C followup
AUC_calculate_48dpfMEANS_38C %>% dplyr::group_by(Family_EBV) %>%  dplyr::summarise(mean = mean(meanAUC.cor.norm))
# High	562.0305			
# Low	  281.1019	
((562.0305-281.1019)/562.0305)*100 # 50% lower metabolic activity on average for the low EBV than High EBV fams



#  TWO WAY ANOVA  ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# run two way anpva for EBV * temperature

TwoWayANOVA_48dpf <- aov(lm(meanAUC.cor.norm ~ Family_EBV*Treatment, data = AUC_calculate_48dpfMEANS))
# check model assumptions 
shapiro.test(resid(TwoWayANOVA_48dpf)) # 0.9866 - pass 
car::leveneTest(TwoWayANOVA_48dpf) # 0.06266 - pass
# conclusion, pass run ANOVA summary 
summary(TwoWayANOVA_48dpf) # significant effect of EBV
#                      Df  Sum Sq Mean Sq F value  Pr(>F)   
# Family_EBV            1  583380  583380  12.524 0.00104 **
# Treatment             3   59315   19772   0.424 0.73651   
# Family_EBV:Treatment  3  291463   97154   2.086 0.11737   
# Residuals            40 1863229   46581 

AUC_calculate_48dpfMEANS %>% dplyr::group_by(Family_EBV) %>%  dplyr::summarise(mean = mean(meanAUC.cor.norm))
# High	551.2879			
# Low	  330.7999
((551.2879-330.7999)/551.2879)*100 # 40% low metabolic activity on average for the low EBV than High EBV fams
```


#### Outputs 
* Plot 
* master file for 72h, 15d, and 48d

```{r Output CalculatedAUC larvae and early seed}

Master_larvae_seed <- rbind(AUC_calculate_72hpf,
                            AUC_calculate_15dpf,
                            AUC_calculate_48dpf)


write.csv(Master_larvae_seed, "output/resazurin/calculated/CalulatedAUC_Master_larvae_earlyseed.csv")

```

```{r Output Resaz_larval_juvenile_MEANse Plot}

Resaz_larval_juvenile_MEANse <- ggpubr::ggarrange(Plot_72hpf_MEAN.SE,
                                                  Plot_15dpf_MEAN.SE,
                                                  Plot_48dpf_MEAN.SE,
                                                  nrow=3)

pdf("output/resazurin/Resaz_larval_juvenile_MEANse.pdf", width = 5, height = 12)
Resaz_larval_juvenile_MEANse
dev.off()

```


## Temperature Ramp


#### Master file


ADJUSTING FOR GAIN, NORMALIZE TO 65
  * Going from lower gain to higher
          AUC * 0.145 for every 1 unit increase in gain 
  * Going from higher gain to lower
          AUC * 0.095 for every 1 unit decrease in gain
          
          
```{r TempRamp clean and parse}

0.095*(77-65)
AUC_calculate_20260702_28C.FINAL <-AUC_calculate_20260630_28C$AUC.df %>% 
            dplyr::filter(gain == 77) %>% 
            dplyr::mutate(AUC.BlankShell = (AUC_calculate_20260630_28C$AUCblanks %>% dplyr::filter(gain == 77 & Type %in% 'shell'))$meanAUC,
                          AUC.cor = AUC - AUC.BlankShell
            )
                          
AUC_calculate_20260702_30C.FINAL <- merge(AUC_calculate_20260702_30C$AUC.df,
                                    (as.data.frame(AUC_calculate_20260702_30C$AUCblanks) %>% dplyr::filter(Type %in% 'shell') 
                                     %>% dplyr::select('gain', 'meanAUC')),by = 'gain') %>% 
                                          dplyr::rename(AUC.BlankShell = meanAUC) %>% 
                                          dplyr::mutate(AUC.cor = AUC - AUC.BlankShell) %>% 
                                          dplyr::filter(gain == 65)
                                  
                     
AUC_calculate_20260704_32C.FINAL <- merge(AUC_calculate_20260704_32C$AUC.df,
                                    (as.data.frame(AUC_calculate_20260704_32C$AUCblanks) %>% dplyr::filter(Type %in% 'shell') 
                                     %>% dplyr::select('gain', 'meanAUC')), by = 'gain') %>% 
                                          dplyr::rename(AUC.BlankShell = meanAUC) %>% 
                                          dplyr::mutate(AUC.cor = AUC - AUC.BlankShell)


AUC_calculate_20260706_34C.FINAL <- merge(AUC_calculate_20260706_34C$AUC.df,
                                    (as.data.frame(AUC_calculate_20260706_34C$AUCblanks) %>% dplyr::filter(Type %in% 'shell') 
                                     %>% dplyr::select('gain', 'meanAUC')), by = 'gain') %>% 
                                          dplyr::rename(AUC.BlankShell = meanAUC) %>% 
                                          dplyr::mutate(AUC.cor = AUC - AUC.BlankShell) %>% 
                                          dplyr::filter(gain == 65)



0.145 * (65-60)
AUC_calculate_20260710_38C.FINAL <- merge(AUC_calculate_20260710_38C$AUC.df,
                                    (as.data.frame(AUC_calculate_20260710_38C$AUCblanks) %>% dplyr::filter(Type %in% 'shell') 
                                     %>% dplyr::select('gain', 'meanAUC')), by = 'gain') %>% 
                                          dplyr::rename(AUC.BlankShell = meanAUC) %>% 
                                          dplyr::mutate(AUC.cor = (AUC - AUC.BlankShell)*1.725) # convert to 65 gain


```

```{r TempRamp output CalculatedAUC}
Master_TempRamp <- rbind(AUC_calculate_20260702_30C.FINAL,
                         AUC_calculate_20260704_32C.FINAL,
                         AUC_calculate_20260706_34C.FINAL,
                         AUC_calculate_20260710_38C.FINAL) %>% 
                  dplyr::mutate(Family_EBV = 
                                  dplyr::case_when(Family > 35 ~ 'Low',
                                                   Family < 35 ~ 'High')
                                            )
nrow(Master_TempRamp)
TempRamp_metadata <- metadata_counts %>% dplyr::select(Treatment, Family, Well, Age, Dry.Weight)

Master_TempRamp_merged <- merge(Master_TempRamp,TempRamp_metadata ) %>% 
                                         dplyr::mutate(AUC.cor.norm = AUC.cor / Dry.Weight)
write.csv(Master_TempRamp_merged, "output/resazurin/calculated/CalulatedAUC_Master_TempRamp.csv")

```

```{r TempRamp plotting}
Master_meansFamily <- Master_merged %>% 
                        dplyr::group_by(Date, Treatment, Family, Age) %>% 
                        dplyr::summarise(meanAUC = mean(AUC.cor.norm)) %>% 
                        dplyr::mutate(Family_EBV = dplyr::case_when(Family >= 48 ~ 'Low',
                                                                    Family <= 20 ~ 'High'))

ggplot(data = Master_meansFamily,
       aes(x = Treatment,
           y = meanAUC,
           color = Family_EBV,
           group = Family)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
                                    theme_classic() 


 ggplot(data =Master_meansFamily, 
                                   aes(x=Treatment, 
                                       y = meanAUC, 
                                        group =  Family_EBV, 
                                        color = Family_EBV)) +
                                    stat_summary(
                                      fun = mean,
                                      geom = "point",
                                      size = 3
                                    ) +
                                    stat_summary(
                                      fun.data = mean_se,
                                      geom = "errorbar",
                                      width = 0.2
                                    ) +
                                    labs(
                                      x = "Treatment",
                                      y = "Area under curve (Resazurin, correct to blank shell)",
                                      title = "C. magallana juvniles temp ramp"
                                    )  +
                                    geom_smooth(method = "loess",               
                                                se = TRUE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                                    theme_classic() 
 
 
 

dodge_width <- 0.70



Master_meansFamily$Family_EBV <- factor(
  Master_meansFamily$Family_EBV,
  levels = c("Low", "High")
)
      
TempRamp_MEAN.SE <- ggplot(
                            Master_meansFamily,
                            aes(
                              x = Treatment,
                              y = meanAUC,
                              colour = Family_EBV,
                              fill = Family_EBV,
                              group = Family_EBV
                            )
                          ) +
                            geom_point(
                              shape = 21,
                              size = 2.4,
                              alpha = 0.55,
                              position = position_jitterdodge(
                                jitter.width = 0.12,
                                dodge.width = dodge_width,
                                seed = 123
                              )
                            ) +
                            scale_fill_manual(values = c('#0072B2', '#E69F00')) +
                            stat_summary(
                              fun = mean,
                              geom = "point",
                              shape = 21,
                              size = 4,
                              colour = "black",
                              position = position_dodge(width = dodge_width)
                            ) +
                            stat_summary(
                              fun.data = mean_se,
                              geom = "errorbar",
                              width = 0.12,
                              linewidth = 0.8,
                              colour = "black",
                              position = position_dodge(width = dodge_width)
                            ) +
                            scale_colour_manual(values = c('#0072B2', '#E69F00')) +
                            labs(
                              title = "Temperature ramp with juveniles",
                              x = "Temperature treatment",
                              y = "AUC (per individual dry weight)",
                              colour = "Independent variable 2",
                              fill = "Independent variable 2"
                            ) +
                            geom_smooth(method = "loess",               
                                                se = FALSE,                      
                                                level = 0.95,
                                                span = 0.8)  +
                            theme_classic() +
                            theme(legend.position = "none")
```

```{r Output Resaz_TempRamp_MEANse Plot}

pdf("output/resazurin/Resaz_TempRamp_MEANse.pdf", width = 6, height = 5)
TempRamp_MEAN.SE
dev.off()

```

* CONCLUSION: Sig effect at 38C (welch's t test) showing 29% lower metabolic activity for High EBV fams than Low EBV fams
              Sig effect of temperature (two way anova) showing 45% lower metabolic activity at 32C than at 38C for all animals measured
```{r stats for Tempramp}

#  T TEST ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# test for differences between  high and low EBV families WITHIN each temperature incubation

# prepare subset to run t tests within temperature
unique(Master_TempRamp_merged$Treatment) # "30C" "32C" "34C" "38C"
TempRamp_MEANS_30C <- Master_TempRamp_merged %>% 
                                      dplyr::filter(Treatment %in% '30C') %>% 
                                      dplyr::group_by(Family, Family_EBV) %>% 
                                      dplyr::summarise(meanAUC.cor.norm = mean(AUC.cor.norm, na.rm = TRUE)) %>% 
                                      dplyr::filter(meanAUC.cor.norm > 0) %>% # omit negatives
                                      na.omit() # omit NAs (those ith no dry weights)
                                      
TempRamp_MEANS_32C <- Master_TempRamp_merged %>% 
                                      dplyr::filter(Treatment %in% '32C') %>% 
                                      dplyr::group_by(Family, Family_EBV) %>% 
                                      dplyr::summarise(meanAUC.cor.norm = mean(AUC.cor.norm, na.rm = TRUE))%>% 
                                      dplyr::filter(meanAUC.cor.norm > 0) %>% # omit negatives
                                      na.omit() # omit NAs (those ith no dry weights)

TempRamp_MEANS_34C <- Master_TempRamp_merged %>% 
                                      dplyr::filter(Treatment %in% '34C') %>% 
                                      dplyr::group_by(Family, Family_EBV) %>% 
                                      dplyr::summarise(meanAUC.cor.norm = mean(AUC.cor.norm, na.rm = TRUE))%>% 
                                      dplyr::filter(meanAUC.cor.norm > 0) %>% # omit negatives
                                      na.omit() # omit NAs (those ith no dry weights)

TempRamp_MEANS_38C <- Master_TempRamp_merged %>% 
                                      dplyr::filter(Treatment %in% '38C') %>% 
                                      dplyr::group_by(Family, Family_EBV) %>% 
                                      dplyr::summarise(meanAUC.cor.norm = mean(AUC.cor.norm, na.rm = TRUE))%>% 
                                      dplyr::filter(meanAUC.cor.norm > 0) %>% # omit negatives
                                      na.omit() # omit NAs (those ith no dry weights)

# MODEL ASSUMPTIONS for t test
# shapiro wilk to check for normality 
by(TempRamp_MEANS_30C$meanAUC.cor.norm, TempRamp_MEANS_30C$Family_EBV, shapiro.test) # 0.4235 - normal
by(TempRamp_MEANS_32C$meanAUC.cor.norm, TempRamp_MEANS_32C$Family_EBV, shapiro.test) # 0.5106 - normal
by(TempRamp_MEANS_34C$meanAUC.cor.norm, TempRamp_MEANS_34C$Family_EBV, shapiro.test) # 0.4167 - normal
by(TempRamp_MEANS_38C$meanAUC.cor.norm, TempRamp_MEANS_38C$Family_EBV, shapiro.test) # 0.7979 - normal

# levenes test to check for homogeneity of variance 
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = TempRamp_MEANS_30C) # 0.4958 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = TempRamp_MEANS_32C) # 0.2348 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = TempRamp_MEANS_34C) # 0.212 - pass
car::leveneTest(meanAUC.cor.norm ~ Family_EBV, data = TempRamp_MEANS_38C) # 0.3857 - pass

# conclusion
# all models for the four temperatures pass! Move forward with t test
t.test(meanAUC.cor.norm ~ Family_EBV, data = TempRamp_MEANS_30C, var.equal = FALSE) # 0.7409 - not significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = TempRamp_MEANS_32C, var.equal = FALSE) # 0.192 -  not significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = TempRamp_MEANS_34C, var.equal = FALSE) # 0.617 - not significant
t.test(meanAUC.cor.norm ~ Family_EBV, data = TempRamp_MEANS_38C, var.equal = FALSE) # 0.04971 - significant

#38C followup
TempRamp_MEANS_38C %>% dplyr::group_by(Family_EBV) %>%  dplyr::summarise(mean = mean(meanAUC.cor.norm))
# High	586.6778			
# Low	  202.7337	
((586.6778-202.7337)/586.6778)*100 # 65% lower metabolic activity on average for the low EBV than High EBV fams


#38C followup
AUC_calculate_48dpfMEANS_38C %>% dplyr::group_by(Family_EBV) %>%  dplyr::summarise(mean = mean(meanAUC.cor.norm))
# High	3071300			
# Low	  4317371	
((4317371-3071300)/4317371)*100 # 29% lower  metabolic activity on average for the High EBV than Low EBV fams



#  TWO WAY ANOVA  ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::
# run two way anpva for EBV * temperature

TempRamp_MEANS <- Master_TempRamp_merged %>%  
                    group_by(Treatment, Family, Family_EBV) %>% 
                    dplyr::summarise(meanAUC.cor.norm = mean(AUC.cor.norm, na.rm = TRUE))%>% 
                    dplyr::filter(meanAUC.cor.norm > 0) %>% # omit negatives
                    na.omit() # omit NAs (those ith no dry weights)
TwoWayANOVA_TempRamp <- aov(lm(meanAUC.cor.norm ~ Family_EBV*Treatment, data = TempRamp_MEANS))
# check model assumptions 
shapiro.test(resid(TwoWayANOVA_TempRamp)) # 0.2932 - pass 
car::leveneTest(TwoWayANOVA_TempRamp) # 0.4986 - pass
# run the ANOVA
summary(TwoWayANOVA_TempRamp) # sig effect of temperature treatment
#                       Df    Sum Sq   Mean Sq F value  Pr(>F)   
# Family_EBV            1 9.878e+08 9.878e+08   0.001 0.97692   
# Treatment             3 1.717e+13 5.723e+12   4.912 0.00556 **
# Family_EBV:Treatment  3 6.342e+12 2.114e+12   1.814 0.16098   
# Residuals            38 4.428e+13 1.165e+12 

TwoWayANOVA_TempRamp_TUKEY <- emmeans::emmeans(TwoWayANOVA_TempRamp, ~ Treatment, type = "response")
letters_table <- multcomp::cld(
                              TwoWayANOVA_TempRamp_TUKEY,
                              adjust = "tukey",
                              Letters = letters
                            )
# Treatment  emmean     SE df lower.CL upper.CL .group
# 32C       2043270 312000 38  1228586  2857953  a    
# 34C       2655614 312000 38  1840930  3470297  ab   
# 30C       3037697 341000 38  2145256  3930138  ab   
# 38C       3694335 312000 38  2879652  4509019   b 
((3694335-2043270)/3694335)*100 #45% lower metabolic activity at 32C than at 38C on average for all animals measured

```
