---
title: "1_mito_stress_analysis"
author: "Samuel Gurr"
date: "2026-03-21"
output: html_document
---

```{r}
library(readxl)
library(stringr)
library(pracma)
library(ggplot2)
library(ggpubr)
library(car)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
knitr::opts_knit$set(root.dir = "C:/Users/gurrs/Documents/6_Github_repositories/GurrLab/Predictive_phenomics/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Airradians_Popgen_OA/") # Sam's
```


```{r load data}

seahorse_25C  <- read.csv("data/Eastern_oyster/seahorse/OCR_Data_norm_AARot_25oC.csv", sep=',', head=T, check.names = FALSE)

seahorse_34C  <- read.csv("data/Eastern_oyster/seahorse/OCR_Data_norm_AARot_34oC.csv", sep=',', head=T, check.names = FALSE)

```

```{r}
names(seahorse_25C)[1] <- "paradigm"
seahorse_25C$paradigm[seahorse_25C$paradigm == ""] <- NA
seahorse_25C$paradigm <- zoo::na.locf(seahorse_25C$paradigm)

seahorse_25C.clean <- janitor::remove_empty(seahorse_25C, which = "cols")
# Remove the dot and everything after it
names(seahorse_25C.clean) <- sub("\\..*", "", names(seahorse_25C.clean))



names(seahorse_34C)[1] <- "paradigm"
seahorse_34C$paradigm[seahorse_34C$paradigm == ""] <- NA
seahorse_34C$paradigm <- zoo::na.locf(seahorse_34C$paradigm)

seahorse_34C.clean <- janitor::remove_empty(seahorse_34C, which = "cols")
# Remove the dot and everything after it
names(seahorse_34C.clean) <- sub("\\..*", "", names(seahorse_34C.clean))


```

```{r shape wide to long}


# 3. Reshape from wide to long and create a tally column
long_data_25C <- seahorse_25C.clean %>%
                    tidyr::pivot_longer(
                      cols = starts_with(c("2", "3")), 
                      names_to = "Family_ID", 
                      values_to = "O2"
                    ) %>%
                    dplyr::rename(Time_min = `Time (minutes)`) %>% 
                    # Create a sequence tally for each repeated group
                    group_by(paradigm, Time_min) %>%
                    mutate(tally = dplyr::row_number(),
                           Family_ID = sub("\\..*", "", Family_ID)) %>% 
                    dplyr::ungroup()


long_data_25C_MeanSE <- Rmisc::summarySE(long_data_25C, measurevar = "O2", groupvars = c("Time_min", "paradigm", "Family_ID"))



# 3. Reshape from wide to long and create a tally column
long_data_34C <- seahorse_34C.clean %>%
                    tidyr::pivot_longer(
                      cols = starts_with(c("2", "3")), 
                      names_to = "Family_ID", 
                      values_to = "O2"
                    ) %>%
                    dplyr::rename(Time_min = `Time (minutes)`) %>% 
                    # Create a sequence tally for each repeated group
                    group_by(paradigm, Time_min) %>%
                    mutate(tally = dplyr::row_number()) %>% 
                    dplyr::ungroup()


long_data_34C <- long_data_34C %>% dplyr::filter(!Family_ID %in% "251") # the only family that does not have any viable data at 25C, omit for comarison

long_data_34C$O2       <- as.numeric(long_data_34C$O2)
long_data_34C$Time_min <- as.numeric(long_data_34C$Time_min)
long_data_34C_MeanSE <- Rmisc::summarySE(long_data_34C, measurevar = "O2", groupvars = c("Time_min", "paradigm", "Family_ID"))

```


```{r view plots 25C}

Plot_MEANSe_25C <- ggplot(long_data_25C, aes(x = Time_min, 
                                             y = O2, 
                                             colour = Family_ID, 
                                             group = Family_ID)) +
                          stat_summary(fun = mean, geom = "line") +
                          stat_summary(fun = mean, geom = "point", size = 2) +
                          stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1)  + 
                          scale_x_continuous(breaks = seq(0, 300, by = 25)) +
                          theme_bw()

# rects_25C <- long_data_25C %>%
#   group_by(paradigm) %>%
#   dplyr::summarize(
#             start = min(Time_min),
#             end = max(Time_min)
#     )
# 
# ggplot() +
#   # 1. Add background shading
#   geom_rect(data = rects_25C, 
#             aes(xmin = start - 0.5, xmax = end + 0.5, 
#                 ymin = -Inf, ymax = Inf, fill = paradigm), 
#             alpha = 0.3) +
#   # Add the mean and SE points/errorbars on top
#   geom_point(
#     data = long_data_25C_MeanSE, 
#     aes(x = Time_min, y = O2, fill = "Family_ID"), 
#     size = 4
#   ) +
#   geom_line(
#     data = long_data_25C_MeanSE, 
#     aes(x = Time_min, y = O2, fill = "Family_ID")) +
#   geom_errorbar(
#     data = long_data_25C_MeanSE, 
#     aes(x = Time_min, ymin = se, ymax = se), 
#     width = 0.2, linewidth = 1, group = "Family_ID"
#   ) +
#   # Formatting and labels
#   theme_classic()

```


```{r view plots 34C}

Plot_MEANSe_34C <- ggplot(long_data_34C, aes(x = Time_min, 
                                             y = O2, 
                                             colour = Family_ID, 
                                             group = Family_ID)) +
                            stat_summary(fun = mean, geom = "line") +
                            stat_summary(fun = mean, geom = "point", size = 2) +
                            stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) + 
                            scale_x_continuous(breaks = seq(0, 300, by = 25)) +
                            theme_bw()

```


```{r}

seahorse_34C.clean


# 3. Reshape from wide to long and create a tally column
long_data_25C.clean <- seahorse_25C.clean %>%
                            tidyr::pivot_longer(
                              cols = starts_with(c("2", "3")), 
                              names_to = "Family_ID", 
                              values_to = "O2"
                            ) %>%
                            # Create a sequence tally for each repeated group
                            group_by(paradigm, `Time (minutes)`) %>%
                            mutate(tally = dplyr::row_number(),
                                   Family_ID = sub("\\..*", "", Family_ID)) %>% 
                            dplyr::ungroup()



# 3. Reshape from wide to long and create a tally column
long_data_34C.clean <- seahorse_34C.clean %>%
                            tidyr::pivot_longer(
                              cols = starts_with(c("2", "3")), 
                              names_to = "Family_ID", 
                              values_to = "O2"
                            ) %>%
                            # Create a sequence tally for each repeated group
                            group_by(paradigm, `Time (minutes)`) %>%
                            mutate(tally = dplyr::row_number(),
                                   Family_ID = sub("\\..*", "", Family_ID)) %>% 
                            dplyr::ungroup()

long_data_34C.clean <- long_data_34C.clean %>% dplyr::filter(!Family_ID %in% "251") # the only family that does not have any viable data at 25C, omit for comarison


```

```{r}

ggarrange(
  Plot_MEANSe_25C,
  Plot_MEANSe_34C,
  nrow = 2
)
```

### Calculations and target values 

  * **Baseline** - initial measurements, once cycles staballize and before  Olgo
  
  * **Maximal respiration** - (after FCCP injection) - (Non-mitochondrial O2 consumption ater Rot/AA) 
  
  * **Spare capacity** - (Maximal respiration) - (Baseline)
  
  * **ATP-linked respiration** - (Basal respiration) - (OCR after Oligo stabalizes)
  
  * **Proton leak** - (OCR after Oligo stabalizes) - (Non-mitochondrial O2 consumption ater Rot/AA) 


```{r Cals 25C}

# Minutes 50-60 =  **Baseline** - initial measurements, three cycles when stabilized BEFORE Olgo

# Minutes 160-175 = **Oligo** three cycles when stabilized AFTER Olgo
 
# Minutes 187-208 * **MaxResp** = Maximal respiration** - (after FCCP injection) - (Non-mitochondrial O2 consumption AFTER Rot/AA) 
 
# Minutes Minutes 258-280 = **NonMito** Non-mitochondrial** O2 consumption AFTER Rot/AA


# * **Spare capacity** - (Maximal respiration) - (Basal respiration) = (MaxResp) - (Baseline)
 
# * **ATP-linked respiration** - (Basal respiration) - (OCR after Oligo stabalizes) = (Oligo) - (Baseline)
 
# * **Proton leak** - (OCR after Oligo stabalizes) - (Non-mitochondrial O2 consumption AFTER Rot/AA)  = (Oligo) - (NonMito)


long_data_25C.clean$`Time (minutes)` = as.numeric(long_data_25C.clean$`Time (minutes)`)
long_data_25C.clean$O2 = as.numeric(long_data_25C.clean$O2)



Baseline_25C = long_data_25C.clean %>% 
                          dplyr::filter(`Time (minutes)` > 49 & `Time (minutes)` < 68) 

Oligo_25C = long_data_25C.clean %>% 
                          dplyr::filter(`Time (minutes)` > 160 & `Time (minutes)` < 178 )

MaxResp_25C = long_data_25C.clean %>% 
                          dplyr::filter(`Time (minutes)` > 187 & `Time (minutes)` < 208 ) 

NonMito_25C = long_data_25C.clean %>% 
                          dplyr::filter(`Time (minutes)` > 258 & `Time (minutes)` < 280)


CalcMaster_25C <- rbind(Baseline_25C,
                        Oligo_25C,
                        MaxResp_25C,
                        NonMito_25C) %>% 
                          dplyr::group_by(paradigm, Family_ID, tally) %>%
                          dplyr::summarise(
                            meanO2 = mean(O2, na.rm = TRUE),
                            seO2  = sd(O2, na.rm = TRUE) / sqrt(dplyr::n()),
                            n = dplyr::n(),
                            .groups  = "drop"
                          )

CalcMaster_25C$meanO2 <- as.numeric(CalcMaster_25C$meanO2)


CalcMaster_25C.final <- CalcMaster_25C %>% 
                  dplyr::select(-c(seO2, n)) %>% 
                  tidyr::pivot_wider(
                              names_from = paradigm,    # Column containing names for the new columns
                              values_from = meanO2     # Column containing values to fill the cells
                            ) %>% 
                  dplyr::mutate(
                    SpareCapacity = (`FCCP 2uM`) - (Baseline),
                    ATPLinked     = (`Oligo 10uM`) - (Baseline),
                    ProtonLeak    = (`Oligo 10uM`) - (`AA/Rot 12/6uM`)
                    ) %>% 
                  dplyr::rename(
                    MaxResp = `FCCP 2uM`,
                    Oligo = `Oligo 10uM`,
                    NonMitoResp = `AA/Rot 12/6uM`
                    ) %>% 
                  tidyr::pivot_longer(
                              cols = -c(Family_ID, tally),
                              names_to = "Metric",    # Column containing names for the new columns
                              values_to = "O2"     # Column containing values to fill the cells
                            ) %>% 
                  dplyr::mutate(Temperature = "25 Celcius")

Plot_Calcs_25C <- ggplot((CalcMaster_25C.final %>% dplyr::filter(!Metric %in% "Oligo")), 
                        aes(x = Family_ID, y = O2, shape = Metric)) +
                        # 1. Plot the group means as points
                        stat_summary(fun = mean, geom = "point", size = 3) +
                        # 2. Add the Standard Error bars automatically
                        stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
                        theme_bw() + 
                        labs(title = "Mito Stress @ 25C") +
                        facet_wrap(~Metric, scales = "free_y")


```


```{r Calcs 34C}
# Minutes 50-60 =  **Baseline** - initial measurements, three cycles when stabilized BEFORE Olgo

# Minutes 160-175 = **Oligo** three cycles when stabilized AFTER Olgo
 
# Minutes 187-208 * **MaxResp** = Maximal respiration** - (after FCCP injection) - (Non-mitochondrial O2 consumption AFTER Rot/AA) 
 
# Minutes Minutes 258-280 = **NonMito** Non-mitochondrial** O2 consumption AFTER Rot/AA


# * **Spare capacity** - (Maximal respiration) - (Basal respiration) = (MaxResp) - (Baseline)
 
# * **ATP-linked respiration** - (Basal respiration) - (OCR after Oligo stabalizes) = (Oligo) - (Baseline)
 
# * **Proton leak** - (OCR after Oligo stabalizes) - (Non-mitochondrial O2 consumption AFTER Rot/AA)  = (Oligo) - (NonMito)


long_data_34C.clean$`Time (minutes)` = as.numeric(long_data_34C.clean$`Time (minutes)`)
long_data_34C.clean$O2 = as.numeric(long_data_34C.clean$O2)



Baseline_34C = long_data_34C.clean %>% 
                          dplyr::filter(`Time (minutes)` > 49 & `Time (minutes)` < 68) 

Oligo_34C = long_data_34C.clean %>% 
                          dplyr::filter(`Time (minutes)` > 160 & `Time (minutes)` < 178 )

MaxResp_34C = long_data_34C.clean %>% 
                          dplyr::filter(`Time (minutes)` > 187 & `Time (minutes)` < 208 ) 

NonMito_34C = long_data_34C.clean %>% 
                          dplyr::filter(`Time (minutes)` > 258 & `Time (minutes)` < 280)


CalcMaster_34C <- rbind(Baseline_34C,
                        Oligo_34C,
                        MaxResp_34C,
                        NonMito_34C) %>% 
                          dplyr::group_by(paradigm, Family_ID, tally) %>%
                          dplyr::summarise(
                            meanO2 = mean(O2, na.rm = TRUE),
                            seO2  = sd(O2, na.rm = TRUE) / sqrt(dplyr::n()),
                            n = dplyr::n(),
                            .groups  = "drop"
                          )

CalcMaster_34C$meanO2 <- as.numeric(CalcMaster_34C$meanO2)


CalcMaster_34C.final <- CalcMaster_34C %>% 
                  dplyr::select(-c(seO2, n)) %>% 
                  tidyr::pivot_wider(
                              names_from = paradigm,    # Column containing names for the new columns
                              values_from = meanO2     # Column containing values to fill the cells
                            ) %>% 
                  dplyr::mutate(
                    SpareCapacity = (`FCCP 2uM`) - (Baseline),
                    ATPLinked     = (`Oligo 10uM`) - (Baseline),
                    ProtonLeak    = (`Oligo 10uM`) - (`AA/Rot 12/6uM`)
                    ) %>% 
                  dplyr::rename(
                    MaxResp = `FCCP 2uM`,
                    Oligo = `Oligo 10uM`,
                    NonMitoResp = `AA/Rot 12/6uM`
                    ) %>% 
                  tidyr::pivot_longer(
                              cols = -c(Family_ID, tally),
                              names_to = "Metric",    # Column containing names for the new columns
                              values_to = "O2"     # Column containing values to fill the cells
                            ) %>% 
                  dplyr::mutate(Temperature = "34 Celcius")

Plot_Calcs_34C <- ggplot((CalcMaster_34C.final %>% dplyr::filter(!Metric %in% "Oligo")), 
                        aes(x = Family_ID, y = O2, shape = Metric)) +
                        # 1. Plot the group means as points
                        stat_summary(fun = mean, geom = "point", size = 3) +
                        # 2. Add the Standard Error bars automatically
                        stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
                        theme_bw() + 
                        labs(title = "Mito Stress @ 34C") +
                        facet_wrap(~Metric, scales = "free_y")


```


```{r}

ggarrange(
          ggarrange(
            Plot_MEANSe_25C,
            Plot_MEANSe_34C,
            nrow = 1,
            ncol = 2
          ), 

          ggarrange(
            Plot_Calcs_25C,
            Plot_Calcs_34C,
            nrow = 1,
            ncol = 2
          ),
          nrow = 2, 
          ncol = 1
)
```
```{r Statistic function}


runanova <- function(datafilename,  outputfilename) {
            # variables for both for loops
            DF_loop           <- data.frame(matrix(nrow = 1, ncol = 12)) # create dataframe to save during for loop
            colnames(DF_loop) <- c('Temperature', 'Metric', 'ShapiroWilk', 'ResidNorm', 
                                   'Levenes', 'HomogVar', 'model', 
                                   'DF.num' , 'DF.denom', 
                                   'F_val','P_val', 'SigDif') # names for comuns in the for loop
            
            unique(datafilename$Groupvar)
            outputfilename <- data.frame()

            for (m in 1:length(unique(datafilename$Metric))) { # call the biodep metrics one by one
              
            DPvar_loop <- unique(as.character(datafilename$Metric))
            Data_DPvar <- datafilename %>% 
                                dplyr::select(Metric, Groupvar, DPvar, INDvar) %>% 
                                dplyr::filter(Metric %in%  DPvar_loop[m])
              
            
            for (i in 1:length(unique(Data_DPvar$Groupvar))) {
              GROUPvar_loop <- unique(Data_DPvar$Groupvar)[i]
              Data_all      <- Data_DPvar %>% dplyr::filter(Groupvar %in% GROUPvar_loop)
              # Data_mean   <- Data_all %>% summarySE(measurevar="DPvar", 
              #                                       groupvars=c("group1","Tank_Replicate", "INDvar")) 
              
              AOVmod      <- aov(lm(Data_all$DPvar ~ as.factor(Data_all$INDvar))) # CHANGE HERE!
              KWmod       <- kruskal.test(Data_all$DPvar  ~ as.factor(Data_all$INDvar)) # CHANGE HERE!
              
              # DF_loop$About <- paste0("Gonad: All bfactor + mean by replicate")
              # DF_loop$About <- paste0("Gonad: All bfactor + all data")  # normality tests for the anova model - asign
              
              
              DF_loop$ShapiroWilk <- shapiro.test(resid(AOVmod))[[2]]
              DF_loop$ResidNorm   <- if( shapiro.test(resid(AOVmod))[[2]] > 0.05) {
                  'YES'} else {'NO'}
              DF_loop$Levenes     <- leveneTest(AOVmod)[[3]][[1]]
              DF_loop$HomogVar    <- if( leveneTest(AOVmod)[[3]][[1]] > 0.05) {
                  'YES'} else {'NO'}
                
              if(shapiro.test(resid(AOVmod))[[2]] > 0.05 & leveneTest(AOVmod)[[3]][[1]] > 0.05) {
                    DF_loop$model       <- 'one-way AOV; x ~ treatment'
                    DF_loop$DF.num      <- summary(AOVmod)[[1]][["Df"]][1]
                    DF_loop$DF.denom    <- summary(AOVmod)[[1]][["Df"]][2]
                    DF_loop$F_val       <- summary(AOVmod)[[1]][["F value"]][1]
                    DF_loop$P_val       <- summary(AOVmod)[[1]][["Pr(>F)"]][1]
                    DF_loop$SigDif      <- if( (summary(AOVmod)[[1]][["Pr(>F)"]][1]) > 0.05) {
                      'NO'} else {'YES'}
            
                  } else {
                    DF_loop$model       <- 'kruskal-wallis; x ~ treatment'
                    DF_loop$DF.num      <- (KWmod)[[2]][["df"]][1]
                    DF_loop$DF.denom    <- NA
                    DF_loop$F_val       <- NA
                    DF_loop$P_val       <- (KWmod)[[3]]
                    DF_loop$SigDif      <- if( ((KWmod)[[3]]) > 0.05) {
                      'NO'} else {'YES'}
                  }
              DF_loop$Temperature     <- GROUPvar_loop
              DF_loop$Metric          <- DPvar_loop[m]
              # asign loop and cumulative output table
              df                      <- data.frame(DF_loop) # name dataframe for this single row
              outputfilename <- rbind(outputfilename,DF_loop) #bind to a cumulative list dataframe
              # print(outputfilename) # print to monitor progress
              
             } # m loop
            } # i loop
return(outputfilename)
}

```


```{r run stats 34C}
MasterFiinal_forStats <- rbind(CalcMaster_25C.final, CalcMaster_34C.final) %>% 
                              dplyr::select(-tally) %>% 
                              dplyr::rename(
                                Groupvar = Temperature, 
                                DPvar  = O2,
                                INDvar   = Family_ID
                              )


runanova(MasterFiinal_forStats, anova_output)
# 34 C significant  for baseline, MAxResp, and ATP linked resp
```


```{r PCA}

# 1. Combine PCA scores with the grouping variable
pca_34C <- data.frame(value = CalcMaster_34C.final$O2, Metric = CalcMaster_34C.final$Metric)

install.packages("data.frame")
?pivot_wider
# FULL NUMERICAL DATASET
data_D1     <- as.data.frame(CalcMaster_34C.final) %>% # call columns , column 1 is Day, 
                           dplyr::mutate(row_id = dplyr::row_number()) %>% 
                           dplyr::select(-c(Temperature, row_id)) %>% #omit columns
                           tidyr::pivot_wider(names_from = Metric,values_from = O2 )
                           reshape2::dcast(row_id ~ Metric, value.var = "O2") 
numerical_data_D1 <- data_D1 %>%  dplyr::select(-c(tally, Family_ID))

familyID_data_D1 <- data_D1 %>%  dplyr::select(Family_ID)

# NORMALIZE DATA
data_normalized_D1     <- scale(numerical_data_D1)


# CORRELATOIN MATRIX AND PLOT IT 
corr_matrix_D1     <- cor(data_normalized_D1)
ggcorrplot::ggcorrplot(corr_matrix_D1)

# RUN PCA 
data.pca_D1 <- princomp(corr_matrix_D1)
summary(data.pca_D1)
# Importance of components:
#                          Comp.1    Comp.2     Comp.3      Comp.4       Comp.5       Comp.6 Comp.7
# Standard deviation     1.311135 0.4107974 0.25603061 0.111272020 8.062230e-09 4.382955e-09      0
# Proportion of Variance 0.874508 0.0858468 0.03334667 0.006298551 3.306581e-17 9.772436e-18      0
# Cumulative Proportion  0.874508 0.9603548 0.99370145 1.000000000 1.000000e+00 1.000000e+00      1

# VIEW PRINCIPLE COMPONENTS 
data.pca_D1$loadings[, 1:2] # of components 1 and 2! 
#                   Comp.1     Comp.2
# NonMitoResp    0.1429498  0.4946832
# Baseline      -0.5205159  0.1299330
# MaxResp       -0.3625646 -0.3421803
# Oligo         -0.3490804 -0.2396364
# SpareCapacity  0.3062096 -0.5952506
# ATPLinked      0.4692816 -0.3217000
# ProtonLeak    -0.3759373 -0.3256975

# we see positibe for ATP and ROS and negative for JC10 and proportion dead 

# EXPLORE LOADINGS 
data.pca_D1_FConly$loadings[, 1:2] # of components 1 and 2! 


# SCREE PLOT 
factoextra::fviz_eig(data.pca_D1, addlabels = TRUE)
# bIPLOT 
factoextra::fviz_pca_var(data.pca_D1, col.var = "black")
# COSIGN CONTRIBUTION 
factoextra::fviz_cos2(data.pca_D1, choice = "var", axes = 1:2) # membrane potential and ROS

# BIPLOT WITH COSIGN 

fviz_pca_var(data.pca_D1, col.var = "cos2",
            gradient.cols = c("green", "orange", "red"),
            repel = TRUE)

pca_res_D1       <- prcomp(data_normalized_D1, scale. = TRUE)
# Merge_master_D1  <- Merge_master_OM_D1 %>% dplyr::mutate(pCO2_all = paste0(pCO2_history, " x ", pCO2_exposure))

# vars for plotting
cols = c('#009E73','#009E73','#009E73',
         '#E69F00','#E69F00','#E69F00',
         '#CC79A7','#CC79A7','#CC79A7')

shapes = c(16, 15, 17,
           16, 15, 17,
           16, 15, 17)


PCAbiplot_D1         <- fviz_pca_biplot(pca_res_D1, 
                          label="var", 
                          habillage=Merge_master_D1$pCO2_all) +
                          ggforce::geom_mark_ellipse(aes(fill = Groups, color = Groups)) +

                          scale_fill_manual(breaks = unique(Merge_master_D1$pCO2_all),
                                          values = cols)+
                          scale_colour_manual(breaks = unique(Merge_master_D1$pCO2_all),
                                          values = cols) +
                          scale_fill_manual(breaks = unique(Merge_master_D1$pCO2_all),
                                          values = cols)+
                          scale_shape_manual(breaks = unique(Merge_master_D1$pCO2_all),
                                          values = shapes) +
                          theme_bw() +
                          xlim(-4, 4) + ylim (-4, 4) +
                          theme(legend.position = 'bottom') +
                          ggtitle("Day 1") +
                          coord_equal() 



PCAbiplot_D1_exp     <- fviz_pca_biplot(pca_res_D1, 
                          label="var", 
                          habillage=familyID_data_D1$Family_ID)
                          # palette = c("#009E73","#E69F00", "#CC79A7")) + 
                          ggforce::geom_mark_ellipse(aes(fill = Groups, color = Groups)) +
                          theme_bw() +
                          scale_shape_manual(breaks = unique(Merge_master_D1$pCO2_exposure),
                                          values = c(16, 15, 17)) +  
                          xlim(-4, 4) + ylim (-4, 4) +
                          theme(legend.position = 'bottom') +
                          ggtitle("Day 1: by exposure") +
                          coord_equal() 

PCAbiplot_D1_history <- fviz_pca_biplot(pca_res_D1, 
                          label="var", 
                          habillage=Merge_master_D1$pCO2_history,
                          palette = c("#009E73","#E69F00", "#CC79A7")) + 
                          ggforce::geom_mark_ellipse(aes(fill = Groups, color = Groups)) +
                          theme_bw() +
                          scale_shape_manual(breaks = unique(Merge_master_D1$pCO2_history),
                                          values = c(16, 15, 17)) +  
                          xlim(-4, 4) + ylim (-4, 4) +
                          theme(legend.position = 'bottom') +
                          ggtitle("Day 1: by history") +
                          coord_equal() 

```

