---
title: "1_mito_stress_analysis"
author: "Samuel Gurr"
date: "2026-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
knitr::opts_knit$set(root.dir = "C:/Users/gurrs/Documents/6_Github_repositories/GurrLab/Predictive_phenomics/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Airradians_Popgen_OA/") # Sam's
```


```{r}
library(readxl)
library(stringr)
library(pracma)
library(ggplot2)
library(ggpubr)
```

```{r load data}

seahorse_25C  <- read.csv("data/Eastern_oyster/seahorse/OCR_Data_norm_AARot_25oC.csv", sep=',', head=T, check.names = FALSE)

seahorse_34C  <- read.csv("data/Eastern_oyster/seahorse/OCR_Data_norm_AARot_34oC.csv", sep=',', head=T, check.names = FALSE)

```

```{r}
names(seahorse_25C)[1] <- "paradigm"
seahorse_25C$paradigm[seahorse_25C$paradigm == ""] <- NA
seahorse_25C$paradigm <- zoo::na.locf(seahorse_25C$paradigm)

seahorse_25C.clean <- janitor::remove_empty(seahorse_25C, which = "cols")
# Remove the dot and everything after it
names(seahorse_25C.clean) <- sub("\\..*", "", names(seahorse_25C.clean))



names(seahorse_34C)[1] <- "paradigm"
seahorse_34C$paradigm[seahorse_34C$paradigm == ""] <- NA
seahorse_34C$paradigm <- zoo::na.locf(seahorse_34C$paradigm)

seahorse_34C.clean <- janitor::remove_empty(seahorse_34C, which = "cols")
# Remove the dot and everything after it
names(seahorse_34C.clean) <- sub("\\..*", "", names(seahorse_34C.clean))


```

```{r shape wide to long}


# 3. Reshape from wide to long and create a tally column
long_data_25C <- seahorse_25C.clean %>%
                    tidyr::pivot_longer(
                      cols = starts_with(c("2", "3")), 
                      names_to = "Family_ID", 
                      values_to = "O2"
                    ) %>%
                    dplyr::rename(Time_min = `Time (minutes)`) %>% 
                    # Create a sequence tally for each repeated group
                    group_by(paradigm, Time_min) %>%
                    mutate(tally = dplyr::row_number(),
                           Family_ID = sub("\\..*", "", Family_ID)) %>% 
                    dplyr::ungroup()


long_data_25C_MeanSE <- Rmisc::summarySE(long_data_25C, measurevar = "O2", groupvars = c("Time_min", "paradigm", "Family_ID"))



# 3. Reshape from wide to long and create a tally column
long_data_34C <- seahorse_34C.clean %>%
                    tidyr::pivot_longer(
                      cols = starts_with(c("2", "3")), 
                      names_to = "Family_ID", 
                      values_to = "O2"
                    ) %>%
                    dplyr::rename(Time_min = `Time (minutes)`) %>% 
                    # Create a sequence tally for each repeated group
                    group_by(paradigm, Time_min) %>%
                    mutate(tally = dplyr::row_number()) %>% 
                    dplyr::ungroup()

long_data_34C$O2       <- as.numeric(long_data_34C$O2)
long_data_34C$Time_min <- as.numeric(long_data_34C$Time_min)
long_data_34C_MeanSE <- Rmisc::summarySE(long_data_34C, measurevar = "O2", groupvars = c("Time_min", "paradigm", "Family_ID"))

```


```{r view plots 25C}

Plot_MEANSe_25C <- ggplot(long_data_25C, aes(x = Time_min, 
                                             y = O2, 
                                             colour = Family_ID, 
                                             group = Family_ID)) +
                          stat_summary(fun = mean, geom = "line") +
                          stat_summary(fun = mean, geom = "point", size = 2) +
                          stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) + 
                          theme_bw()

rects_25C <- long_data_25C %>%
  group_by(paradigm) %>%
  dplyr::summarize(
            start = min(Time_min),
            end = max(Time_min)
    )

ggplot() +
  # 1. Add background shading
  geom_rect(data = rects_25C, 
            aes(xmin = start - 0.5, xmax = end + 0.5, 
                ymin = -Inf, ymax = Inf, fill = paradigm), 
            alpha = 0.3) +
  # Add the mean and SE points/errorbars on top
  geom_point(
    data = long_data_25C_MeanSE, 
    aes(x = Time_min, y = O2, fill = "Family_ID"), 
    size = 4
  ) +
  geom_line(
    data = long_data_25C_MeanSE, 
    aes(x = Time_min, y = O2, fill = "Family_ID")) +
  geom_errorbar(
    data = long_data_25C_MeanSE, 
    aes(x = Time_min, ymin = se, ymax = se), 
    width = 0.2, linewidth = 1, group = "Family_ID"
  ) +
  # Formatting and labels
  theme_classic()

```


```{r view plots 34C}

Plot_MEANSe_34C <- ggplot(long_data_34C, aes(x = Time_min, 
                                             y = O2, 
                                             colour = Family_ID, 
                                             group = Family_ID)) +
                            stat_summary(fun = mean, geom = "line") +
                            stat_summary(fun = mean, geom = "point", size = 2) +
                            stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) + 
                            theme_bw()

rects_34C <- long_data_34C %>%
  group_by(paradigm) %>%
  dplyr::summarize(
            start = min(Time_min),
            end = max(Time_min)
    )

ggplot() +
  # 1. Add background shading
  geom_rect(data = rects_34C, 
            aes(xmin = start - 0.5, xmax = end + 0.5, 
                ymin = -Inf, ymax = Inf, fill = paradigm), 
            alpha = 0.3) +
  # Add the mean and SE points/errorbars on top
  geom_point(
    data = long_data_34C_MeanSE, 
    aes(x = Time_min, y = O2, fill = "Family_ID"), 
    size = 4
  ) +
  geom_line(
    data = long_data_34C_MeanSE, 
    aes(x = Time_min, y = O2, fill = "Family_ID")) +
  geom_errorbar(
    data = long_data_34C_MeanSE, 
    aes(x = Time_min, ymin = se, ymax = se), 
    width = 0.2, linewidth = 1, group = "Family_ID"
  ) +
  # Formatting and labels
  theme_classic()

```

```{r}

seahorse_34C.clean


# 3. Reshape from wide to long and create a tally column
long_data <- seahorse_34C.clean %>%
                    tidyr::pivot_longer(
                      cols = starts_with(c("2", "3")), 
                      names_to = "Family_ID", 
                      values_to = "O2"
                    ) %>%
                    # Create a sequence tally for each repeated group
                    group_by(paradigm, `Time (minutes)`) %>%
                    mutate(tally = dplyr::row_number(),
                           Family_ID = sub("\\..*", "", Family_ID)) %>% 
                    dplyr::ungroup()
```

```{r}

ggarrange(
  Plot_MEANSe_25C,
  Plot_MEANSe_34C,
  nrow = 2
)
```

### Calculations and target values 

  * **Baseline** - initial measurements, once cycles staballize and before  Olgo
  
  * **Maximal respiration** - (after FCCP injection) - (Non-mitochondrial O2 consumption ater Rot/AA) 
  
  * **Spare capacity** - (Maximal respiration) - (Baseline)
  
  * **ATP-linked respiration** - (Basal respiration) - (OCR after Oligo stabalizes)
  
  * **Proton leak** - (OCR after Oligo stabalizes) - (Non-mitochondrial O2 consumption ater Rot/AA) 

```{r}


```

```{r}


```

```{r}


```